---
title: Profile of patients triaged green in the Emergency Department (ED) of a secondary care hospital in Mumbai
author: "Anna"
date: "12/05/2020"
output:
  word_document: default
  html_document: 
    fig_caption: yes
    fig_height: 3
    fig_width: 5
---


```{r setup, include=FALSE, results='asis'}
library(knitr)
library(psych)
library(rio)
library(data.table)
library(epiDisplay)
opts_chunk$set(echo = FALSE)
```

# Abstract

## Introduction

Injuries are one of the leading causes of Emergency Department visits. Physicians use localized screening methods to triage patients into Red, Orange, Yellow and Green, from most to least urgent. Screening methods depend on various factors, such as, physiological parameters of the patient, injury severity, anticipated use and availability of resources (incase of mass casualty). The proportion of patients with life-threatning injuries or injuries that require urgent attention is far less in comparison to patients that are triaged green. While, retrospectively considering management of injuried patients, most attention is given to patients that have injuries which may require hospitalization. Thereby disregarding the pool of patients where most of the ED resorurces are spent. Understanding the characteristic of these patients might give us an insight in managing them, inadvertently making effective use of the resources without affecting the quality of care provided.

## Aim

The aim of this study is to describe the demographic, physiological and injury characteristics of trauma patients triaged green in the ED of a secondary healthcare hospital in Mumbai.

## Methods

Data of patients triaged green were extracted from the Trauma Triage Study in India dataset. A retrospective analysis of the data was performed using R. 

## Results
```{r}
# read the original csv file

g <- import("C:\\Users\\annan\\OneDrive\\Documents\\Projects\\PGTP\\ttris-dataset-with-iss-8317-20190708120045.csv")

# Created a dataframe of triage category green

kbbh <- subset(x=g,subset = g$tc == 0)
```

```{r}

pts <- as.numeric(nrow(kbbh))

kbbh$age <- replace(kbbh$age, kbbh$age == 999, NA)

mean_age <- round(mean(kbbh$age, na.rm = TRUE),2)

stddev_age <- round(sd(kbbh$age, na.rm = TRUE),2)

```


In the TTRIS dataset, `r pts` patients were triaged green. Mean age of these patients was `r mean_age` (+-`r stddev_age`). 97% of the patients were below 65 years of age. 76.7% patients were male, while 23.3% patients were female. 98.5% patients had blunt injuries.Most common mode of arrival to the ED was through private vehicles (80.1%). 13.6% patients came to the ED on foot.
Patients with history of assualt were the highest followed by transport accidents. Injury due to animal bites were significantly high. 

```{r}

# Categorizing ISS Score

kbbh$isspv[kbbh$iss >= 9 & kbbh$iss <= 15] = 'Moderate'
kbbh$isspv[kbbh$iss == 1 & kbbh$iss <= 8] = 'Minor'
kbbh$isspv[is.na(kbbh$iss)] = 'No Injuries defined'

Table1 <- xtabs(~kbbh$isspv, data = kbbh)
kable(Table1, 
      caption = "Table1. ISS Score of patients",
)

```

```{r}
# For calculating RTS

## Append a new variable 'GCS' to the dataset

kbbh$GCS <- kbbh$egcs + kbbh$vgcs + kbbh$mgcs

kbbh$GCS <- replace(kbbh$GCS, kbbh$GCS > 15, NA)

kbbh$GCSpv[kbbh$GCS==15|kbbh$GCS==14|kbbh$GCS==13] = 4
kbbh$GCSpv[kbbh$GCS==12|kbbh$GCS==11|kbbh$GCS==10|kbbh$GCS==9] = 3
kbbh$GCSpv[kbbh$GCS==8|kbbh$GCS==7|kbbh$GCS==6] = 2
kbbh$GCSpv[kbbh$GCS==5|kbbh$GCS==4] = 1
kbbh$GCSpv[kbbh$GCS<4] = 0

## RR parameter value for calculating RTS

kbbh$rr <- replace(kbbh$rr, kbbh$rr == 999, NA)

kbbh$rrpv[kbbh$rr >= 10 & kbbh$rr <= 29] = 4
kbbh$rrpv[kbbh$rr > 29] = 3
kbbh$rrpv[kbbh$rr >= 6 & kbbh$rr <= 9] = 2
kbbh$rrpv[kbbh$rr >= 1 & kbbh$rr <= 5] = 1
kbbh$rrpv[kbbh$rr == 0] = 0

## SBP parameter value for calculating RTS

kbbh$sbp <- replace(kbbh$sbp, kbbh$sbp == 999, NA)

kbbh$sbppv[kbbh$sbp > 89] = 4
kbbh$sbppv[kbbh$sbp >= 76 & kbbh$sbp <= 89] = 3
kbbh$sbppv[kbbh$sbp >= 50 & kbbh$sbp <= 75] = 2
kbbh$sbppv[kbbh$sbp >= 1 & kbbh$sbp <= 49] = 1
kbbh$sbppv[kbbh$sbp == 0] = 0

## Formula for calculating RTS

kbbh$RTS = round(0.9368*as.numeric(kbbh$GCSpv) + 
                   0.7326*as.numeric(kbbh$sbppv) + 
                   0.2908*as.numeric(kbbh$rrpv), digits = 0)

```

```{r}
Table2 <- xtabs(~kbbh$isspv+kbbh$RTS, data = kbbh)
kable(Table2, 
      caption = "Table2. ISS and RTS of patients",
)
```

## Conclusion

Among the patients triaged green, injury due to assualt and transport accidents were predominant.More than half of the patients had no injuries defined with an RTS score of 8. 

## Analysis

```{r}
# Cateogrising Age

## Create age groups
agebreaks <- c(17,25,45,65,98)
agelabels <- c("18-24","25-44","45-64","65+")
setDT(kbbh)[,agegrp := cut(kbbh$age,
                           breaks = agebreaks,
                           right = FALSE,
                           labels = agelabels)]

## Table of age group
tab1(kbbh$agegrp, main = "Distribution of patients by Age")

```

```{r}

# Table of Sex

kbbh$sex <- replace(kbbh$sex, kbbh$sex == 999, NA)

tab1(kbbh$sex, main = "Distribution of patients by Gender")

```

```{r}
kbbh$tyi <- replace(kbbh$tyi, kbbh$tyi == 999, NA)

tab1(kbbh$tyi, main = "Distribution of type of injury")

```

```{r}
kbbh$mot <- replace(kbbh$mot, kbbh$mot == 999, NA)

tab1(kbbh$mot, main = "Distribution of Mode of transport")

```

```{r}

tab1(kbbh$RTS, main = "Distribution of RTS")

```

## Categorizing moi

### 1
```{r}

library(dplyr)

kbbh$moi2 <- mutate(kbbh,
                    moi = case_when(
                      moi %in% V010:V99 ~ "Transport Accidents"))

     
table(kbbh$moi)
```

```{r}
### 2

library(tidyverse)
library(stringr)
library(forcats)

assign_moi  <- function(code){
  if(str_sub(code, 2) %in% as.factor(c(
    "V1","V2","V3","V4","V5","V6","V7","V8","V9")){
    "Transport Accidents"
  } else {
      code
  }
}

kbbh %>% mutate(moi = fct_relabel(moi, assign_moi))

class(kbbh$moi)
```

```{r}

### 3

ifelse(kbbh$moi %in% c("V01","V010","V011","V019","V02",
                  "V020","V021","V029","V03","V030","V031","V039",
                  "V099"), return("Accidents"),return("YTBD"))


```



```{r}
### 4

accidents <- c("V01","V010","V011","V019","V02",
               "V020","V021","V029","V03","V030","V031","V039",
               "V099")

moi <- transform(kbbh$moi, newcat = 
                   ifelse(
                     moi %in% accidents,
                     "yes!",
                     "no"
                   ))

```

