---
title: Profile of patients triaged green in the Emergency Department (ED) of a secondary care hospital in Mumbai
author: "Anna"
date: "12/05/2020"
output:
  word_document: default
  html_document: 
    fig_caption: yes
    fig_height: 3
    fig_width: 5
---


```{r setup, include=FALSE, results='asis', message=FALSE}
library(knitr)
library(dplyr)
library(psych)
library(data.table)
library(tableone)
library(epiDisplay)
library(tigerstats)
opts_chunk$set(echo = FALSE)
```

# Abstract

## Introduction

Injuries are one of the leading causes of Emergency Department visits.
Physicians triage patients into Red, Orange, Yellow and Green, from most to least urgent based on injruy severity and physiological parameters. The proportion of patients that are triaged green is more in comparison to patients with life-threatning injuries and injuries that require urgent attention. It is important to understand the characteristics of these patients to know if there is a need for them to be managed differently so that the manpower and resources of the emergency department can be directed towards serious patients.

## Aim

The aim of this study is to describe the demographic, physiological and injury characteristics of trauma patients triaged green in the ED of a secondary healthcare hospital in Mumbai.

## Methods

Data of patients triaged green were extracted from the Trauma Triage Study in India dataset. A retrospective cohort analysis of the data was performed using R. 

## Results
```{r, warning=FALSE, results='hide'}

# read the original csv file

## datatable with ISS
data.dir <- "C:\\Users\\annan\\OneDrive\\Documents\\Projects\\PGTP\\"
g <- data.table::fread(file.path(data.dir, "ttris-dataset-with-iss-8317-20190708120045.csv"))

## datatable without ISS
data.dir <- "C:\\Users\\annan\\OneDrive\\Documents\\Projects\\PGTP\\"
h <- data.table::fread(file.path(data.dir, "TTRIS.csv"))

#h$iss <- NA

#h[] <- mapply(FUN = as,h,sapply(g,class),SIMPLIFY = FALSE)


## Merged the two datatables

h$pid <- as.integer(h$pid)

merged <- merge(g, h, by = "pid")

## Changed column names

names(merged)[names(merged) == "iss.y"] <- "delcolumn.x"
names(merged)[names(merged) == "iss.x"] <- "iss.y"

## Removing duplicated columns
dt <- data.table(merged)
drop.cols <- grep(".x$", colnames(dt))
dt[, (drop.cols) := NULL]

## Removed .y from the coulmn names

names(dt) <- sub(".y$", "", names(dt))
make.names(names(dt))

```


```{r}

# Created a dataframe of triage category green
kbbh <- subset(x=dt,subset = dt$tc == 0)

# Number of records
dm <- as.numeric(nrow(kbbh))

# Age

kbbh$age <- as.numeric(kbbh$age)

kbbh$age <- replace(kbbh$age, kbbh$age == 999, NA)

mean_age <- round(mean(kbbh$age, na.rm = TRUE),2)

stddev_age <- round(sd(kbbh$age, na.rm = TRUE),2)

## Creating age groups

agebreaks <- c(17,25,45,65,98)
agelabels <- c("18-24","25-44","45-64","65+")
setDT(kbbh)[,agegrp := cut(kbbh$age,
                           breaks = agebreaks,
                           right = FALSE,
                           labels = agelabels)]

## Calculating percentage of patients below age 65

#dm_age <- nrow(subset(kbbh, !is.na(age))) 

age_below65 <- dplyr::filter(kbbh, kbbh$agegrp != "65+") 
nm_age <- nrow(age_below65) # neumerator

percent_agebelow65 <- round((nm_age/ dm) * 100, 2)

# Sex

#dm_sex <- nrow(subset(kbbh, !is.na(sex))) # denominator if excluding NA

kbbh$sex <- as.numeric(kbbh$sex)

male <- dplyr::filter(kbbh, kbbh$sex == 1) # neumerator
nm_male <- nrow(male)
percent_male <- round((nm_male/ dm) * 100,2)

female <- dplyr::filter(kbbh, kbbh$sex == 0) # neumerator
nm_female <- nrow(female)
percent_female <- round((nm_female/ dm) * 100,2)

# Type of Injury

#dm_tyi <- nrow(subset(kbbh, !is.na(tyi))) #denominator if excuding NA

kbbh$tyi <- as.numeric(kbbh$tyi)

blunt <- dplyr::filter(kbbh, kbbh$tyi == 0) # neumerator
nm_blunt <- nrow(blunt)
percent_blunt <- round((nm_blunt/ dm) * 100,2)

# Mode of Transport

kbbh$mot <- as.numeric(kbbh$mot)

kbbh$mot <- replace(kbbh$mot, kbbh$mot == 999, NA)

mot_pvt <- dplyr::filter(kbbh, kbbh$mot == 2) # neumerator
nm_pvt <- nrow(mot_pvt)
percent_mot_pvt <- round((nm_pvt/ dm) * 100,2)

mot_onfoot <- dplyr::filter(kbbh, kbbh$mot == 3) # neumerator
nm_onfoot <- nrow(mot_onfoot)
percent_mot_onfoot <- round((nm_onfoot/ dm) * 100,2)

```


```{r, results='hide'}

# Mechanism of Injury

moi.collapsed <- moi <- as.character(kbbh$moi)

## Method 1

### So, any V will be a transport accident

moi.collapsed[grep("^V[0-9]*$", moi)] <- "Transport accident"

### W00 to W19 are falls
icd.range <- paste0(paste0("^W", stringr::str_pad(0:19, 2, pad = "0"), "$"), collapse = "|")
moi.collapsed[grep(icd.range, moi)] <- "Fall"

### W50 to W64 are animal bites

icd.range <- paste0(paste0("^W", stringr::str_pad(50:64, 2, pad = "0"), "$"), collapse = "|")
moi.collapsed[grep(icd.range, moi)] <- "Animal bite"

## Method 2

## You can probably see how you could create a function to do some work for you
get_icd_range <- function(first.char, start.number, end.number) {
    icd.range <- paste0(
        paste0(
            "^",
            first.char,
            stringr::str_pad(start.number:end.number, 2, pad = "0"),
            "[0-9]?$"),
        collapse = "|")
    return (icd.range)
}

## X85 to X99 and Y00 to Y09 are assault - by using the function created.

icd.range <- paste0(get_icd_range("X", 85, 99), "|", get_icd_range("Y", 0, 9))
moi.collapsed[grep(icd.range, moi)] <- "Assault"

## By now you notice that you repeat moi.collapsed[grep(icd.range,
## moi)] multiple times. You can do better with a loop.

icd.ranges <- list(c("Transport accident" = get_icd_range("V", 0, 99)),
                   c("Fall" = get_icd_range("W", 0, 19)),
                   c("Animal bite" = get_icd_range("W", 50, 64)),
                   c("Assault" = paste0(get_icd_range("X", 85, 99), "|", get_icd_range("Y", 0, 9))))

moi.collapsed2 <- moi ## Just for fun, so that we can compare moi.collapsed and moi.collapsed2

for(icd.range in icd.ranges) moi.collapsed2[grep(icd.range, moi)] <- names(icd.range)

## Check that moi.collapsed and moi.collapsed2 are identical
identical(moi.collapsed, moi.collapsed2) # Should return TRUE
## This means that you can use the function defined as get_icd_range
## together with a list to assign the new values. If you want to
## define all that are not in the list as for example "Other" then you
## can do:

moi.collapsed2[!(moi.collapsed2 %in% sapply(icd.ranges, attr, "names"))] <- "Other"

## And if you want to put this back into the data as a factor then:
kbbh$moi2 <- as.factor(moi.collapsed2)

moi_assualt <- dplyr::filter(kbbh, kbbh$moi2 == "Assault") # neumerator
nm_moi_assualt <- nrow(moi_assualt)
percent_moi_assualt <- round((nm_moi_assualt/ dm) * 100,2)

moi_tra <- dplyr::filter(kbbh, kbbh$moi2 == "Transport accident") # neumerator
nm_moi_tra <- nrow(moi_tra)
percent_moi_tra <- round((nm_moi_tra/ dm) * 100,2)

moi_fall <- dplyr::filter(kbbh, kbbh$moi2 == "Fall") # neumerator
nm_moi_fall <- nrow(moi_fall)
percent_moi_fall <- round((nm_moi_fall/ dm) * 100,2)

moi_anb <- dplyr::filter(kbbh, kbbh$moi2 == "Animal bite") # neumerator
nm_moi_anb <- nrow(moi_anb)
percent_moi_anb <- round((nm_moi_anb/ dm) * 100,2)

moi_other <- dplyr::filter(kbbh, kbbh$moi2 == "Other") # neumerator
nm_moi_other <- nrow(moi_other)
percent_moi_other <- round((nm_moi_other/ dm) * 100,2)

# Trauma outcomes

kbbh$s24h <- as.numeric(kbbh$s24h)
kbbh$s30d <- as.numeric(kbbh$s30d)
kbbh$s6m <- as.numeric(kbbh$s6m)

kbbh$s24h <- replace(kbbh$s24h, kbbh$s24h == 999, NA)
kbbh$s30d <- replace(kbbh$s30d, kbbh$s30d == 999, NA)
kbbh$s6m <- replace(kbbh$s6m, kbbh$s6m == 999, NA)


## Alive at 6month
outcome_a <- dplyr::filter(kbbh, kbbh$s6m == "0") # neumerator
nm_outcome_a <- nrow(outcome_a)
percent_outcome_a <- round((nm_outcome_a/ dm) * 100,2)

## Status unknown/ Loss of followup
outcome_uk <- dplyr::filter(kbbh, is.na(kbbh$s6m)) # neumerator
nm_outcome_uk <- nrow(outcome_uk)
percent_outcome_uk <- round((nm_outcome_uk/ dm) * 100,2)

## Dead at 6month
outcome_d <- dplyr::filter(kbbh, kbbh$s6m == "1") # neumerator
nm_outcome_d <- nrow(outcome_d)
percent_outcome_d <- round((nm_outcome_d/ dm) * 100,2)

## Successful Followups
percent_followup <- percent_outcome_a+percent_outcome_d

```

In the TTRIS dataset, `r dm` patients were triaged green. The mean age of these patients was `r mean_age` (+-`r stddev_age`). `r percent_agebelow65`% of the patients were below 65 years of age. `r percent_male`% patients were male, while `r percent_female`% patients were female. `r percent_blunt`% patients had blunt injuries. Most common mode of arrival to the ED was through private vehicles (`r percent_mot_pvt`%). `r percent_mot_onfoot`% patients came to the ED on foot. Assualt were the most common cause of injury `r percent_moi_assualt`% followed by transport accidents `r percent_moi_tra`% and fall `r percent_moi_fall`. Injury due to animal bites were `r percent_moi_anb`%.`r percent_followup`% patients were successfully followed up at 6months, of which `r nm_outcome_d` patients were reported dead. 

```{r}

kbbh$iss <- as.numeric(kbbh$iss)

# Categorizing ISS Score

kbbh$isspv[kbbh$iss >= 9 & kbbh$iss <= 15] = 'Moderate'
kbbh$isspv[kbbh$iss == 1 & kbbh$iss <= 8] = 'Minor'
kbbh$isspv[is.na(kbbh$iss)] = 'No injuries defined'

Table2 <- colPerc(xtabs(~kbbh$isspv+kbbh$moi2, data = kbbh))
kable(Table2, 
      caption = "Table 1. Percentage distribution of Injury severity among different Mechanism of Injury",
)

```

```{r}
# For calculating RTS

## Append a new variable 'GCS' to the dataset

kbbh$egcs <- as.numeric(kbbh$egcs)
kbbh$vgcs <- as.numeric(kbbh$vgcs)
kbbh$mgcs <- as.numeric(kbbh$mgcs)

gcs <- with(kbbh, egcs + vgcs + mgcs)

gcs <- replace(gcs, gcs > 15, NA)

gcspv <- gcs
gcspv[gcs >= 13] <- 4
gcspv[gcs >= 9 & gcs < 13] <- 3
gcspv[gcs >= 6 & gcs < 9] <- 2
gcspv[gcs >= 4 & gcs < 6] <- 1
gcspv[gcs < 4] <- 0

## RR parameter value for calculating RTS

kbbh$rr <- as.numeric(kbbh$rr)

rr <- with(kbbh,replace(rr, rr == 999, NA))

rrpv <- rr

rrpv[rr >= 10 & rr <= 29] = 4
rrpv[rr > 29] = 3
rrpv[rr >= 6 & rr <= 9] = 2
rrpv[rr >= 1 & rr <= 5] = 1
rrpv[rr == 0] = 0

## SBP parameter value for calculating RTS

kbbh$sbp <- as.numeric(kbbh$sbp)

sbp <- with(kbbh,replace(sbp, sbp == 999, NA))

sbppv <- sbp
sbppv[sbp > 89] <- 4
sbppv[sbp >=76 & sbp <= 89] <- 3
sbppv[sbp >= 50 & sbp <= 75] <- 2
sbppv[sbp >= 1 & sbp <= 49] <- 1
sbppv[sbp == 0] <- 0

## Formula for calculating RTS

## Why do you want to round it using 0 digits?
## So that I get a whole number.The decimals make a longer table.

kbbh$RTS = round(0.9368*as.numeric(gcspv) + 
                 0.7326*as.numeric(sbppv) + 
                 0.2908*as.numeric(rrpv), digits = 0)

Table2 <- rowPerc(xtabs(~kbbh$RTS, data = kbbh))
kable(Table2, 
      caption = "Table 1. RTS of patients",
)

```

## Conclusion

Among the patients triaged green, injury due to assualt and transport accidents were predominant.98.94% patients had an RTS score of 8. Patients with minor injuries and no injuries defined are upto 99.75%. 

## Analysis

```{r, warning=FALSE, results='hide'}

## Table of age group
tab1(kbbh$agegrp, main = "Distribution of patients by Age")

tab1(kbbh$sex, main = "Distribution of patients by Gender")

kbbh$tyi <- replace(kbbh$tyi, kbbh$tyi == 999, NA)

tab1(kbbh$tyi, main = "Distribution of type of injury")

kbbh$mot <- replace(kbbh$mot, kbbh$mot == 999, NA)

tab1(kbbh$mot, main = "Distribution of Mode of transport")

tab1(kbbh$RTS, main = "Distribution of RTS")

tab1(kbbh$hd, main = "Distribution of hospital disposition")

#extra

```



