---
title: "Comparision of different trauma scoring system in adult tbi patients"
author: "Debojit Basak"
date: "02/06/2021"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---
# Abstract
Trauma is one of the leading causes of death in India. Road traffic accidents cause about 60% of all traumatic brain injuries (TBI) in India. About 50% of trauma deaths In India are caused by TBI. The aim of this study is to compare GCS with other trauma scoring systemsin TBI patients.
## Method
Analysis of the Towards Improved Trauma Care Outcomes (TITCO)cohort in India was done. This data was collected from four tertiary care hospitals in India between June 2013 and March 2015. All patients who were 18 years or older and admitted from the emergency department (ER) withTBI were included. The data was analysed to measure the usefulness of each score and also compare the usefulness of the different scoring systems. The area under the Receiver Operating Characteristic (ROC) curve analysis was done for discrimination of the scoring systems.


```{r setup, include=FALSE}
## Load packages
library(rio)
library(tidyr)
library(dplyr)
library(stringr)
library(tableone)
library(pROC)
library(lubridate)
library(DiagrammeR)
library(gtsummary)

```

```{r , adding variables and others}
## Loading the csv file
url <- "https://raw.githubusercontent.com/titco/titco-I/master/titco-I-full-dataset-v1.csv"
titco <- import(url) %>% as_tibble()
nrow(titco)
## Apply filter to find all TBI patients in dataset using ICD 10 codes
icd.codes <- c("s02.0", "s02.1", "s02.8", "s02.9", 
               "s06.0", "s06.1", "s06.2", "s06.3", "s06.4", "s07.1",
               "s06.5", "s06.6", "s06.7", "s06.8", "s06.9", "s08.8",
               "s09.0", "s09.7", "s09.9", "s03.5")
pattern <- str_c(icd.codes, collapse = "|")
titco <- unite(titco, allinj, contains("_icd"), sep = ",", remove = FALSE)

## Filter adult tbi patients
titcotbi <- filter(titco, str_detect(allinj, pattern))
nrow(titcotbi)
## Creating new data set by filtering adults
titcotbi$age <- as.numeric(titcotbi$age)
ptbi <- filter(titcotbi, age < 18 )
nrow(ptbi)
adtbi1 <- filter(titcotbi, age > 17)
#sample size
nrow(adtbi1)

# removing data with incomplete data
adtbi2 <- adtbi1[!is.na(adtbi1$sbp_1),]
adtbi3 <- adtbi2[!is.na(adtbi2$rr_1),]
adtbi4 <- adtbi3[!is.na(adtbi3$gcs_t_1),]
adtbi5 <- adtbi4[!is.na(adtbi4$ti),]


## Trauma scoring system
## Removing NA's for ROC curve of GCS
## You shouldn't remove values in individual variables because that
## will mess up the ordering
## MGAP score
## Type of injury
ti.mgap  <- with(adtbi5, ifelse(ti == "Blunt", 5,
                         ifelse(ti == "Penetrating", 0, NA)))
ti.mgap <- as.numeric(ti.mgap)
summary(ti.mgap)
## Age
age.mgap <- with(adtbi5, ifelse(age >= 60, 0,
                        ifelse(age < 60, 5, NA)))
## Systolic BP
sbp.mgap <- with(adtbi5,
                 ifelse(sbp_1 >= 120, 5,
                 ifelse(sbp_1 >= 60 & sbp_1 < 120, 3,
                 ifelse(sbp_1 < 60,0 ,NA))))
## GCS
gcs.mgap <- adtbi5$gcs_t_1
## Create mgap
adtbi5["mgap"]<- mgap <- ti.mgap + gcs.mgap + age.mgap + sbp.mgap


## GAP Score
## age
age.gap <- with(adtbi5, ifelse(age >= 60, 0,
                       ifelse(age < 60,3, NA)))
## Systolic BP
sbp.gap <- with(adtbi5, ifelse(sbp_1 >= 120, 6,
                       ifelse(sbp_1 >= 60 & sbp_1 < 120, 4,
                       ifelse(sbp_1 < 60, 0, NA))))
## GCS 
gcs.gap <- adtbi5$gcs_t_1
## Create GAP
adtbi5["gap"] <- gap <-  gcs.gap + age.gap  + sbp.gap
summary(gap)
## RTS
# GCS
gcs.rts <- with(adtbi5, ifelse(gcs_t_1 >= 13, 4,
                       ifelse(gcs_t_1 >= 9 & gcs_t_1 < 13, 3,
                       ifelse(gcs_t_1 >= 6 & gcs_t_1 < 9, 2,
                       ifelse(gcs_t_1 >= 4 & gcs_t_1 < 6,1,
                       ifelse(gcs_t_1 < 4, 0, NA))))))
## RR
rr.rts <- with(adtbi5, ifelse(rr_1 >= 10 & rr_1 <= 29, 4,
                      ifelse(rr_1 > 29, 3,
                      ifelse(rr_1 >= 6 & rr_1 <= 9, 2,
                      ifelse(rr_1 >= 1 & rr_1 <= 5,1,
                      ifelse(rr_1 == 0, 0, NA))))))
## SBP
sbp.rts <- with(adtbi5, ifelse(sbp_1 > 89, 4,
                       ifelse(sbp_1 >=76 & sbp_1 <= 89, 3,
                       ifelse(sbp_1 >= 50 & sbp_1 <= 75, 2,
                       ifelse(sbp_1 >= 1 & sbp_1 <= 49,1,
                       ifelse(sbp_1 == 0, 0, NA))))))
## Calculate RTS
rts  <-adtbi5["rts"] <- 0.9368 * gcs.rts + 0.7326 * sbp.rts + 0.2908 * rr.rts
rts <- rts[!is.na(rts)]
summary(rts)
## KTS
## Age
age.kts  <- with(adtbi5, ifelse(age <= 55, 2,
                        ifelse(age > 55, 1, NA)))
## sbp
sbp.kts <- with(adtbi5, ifelse(sbp_1 > 89, 4,
                       ifelse(sbp_1 >=50 & sbp_1 <= 89, 3,
                       ifelse(sbp_1 >= 1 & sbp_1 <= 49, 2,
                       ifelse(sbp_1 == 0, 1, NA)))))
## RR
rr.kts <- with(adtbi5, ifelse(rr_1 >= 10 & rr_1 <= 29 , 3,
                      ifelse(rr_1 >= 30, 2,
                      ifelse(rr_1 <= 9, 1, NA))))
## AVPU
avpu.kts <- adtbi5["AVPU"]<- with(adtbi5, ifelse(gcs_e_1 == 4, 4,
                                        ifelse(gcs_e_1 == 3,3,
                                        ifelse(gcs_e_1 == 2,2,
                                        ifelse(gcs_e_1== 1,1 , NA)))))
## No of serious injuries (nsi)
ais <- select(adtbi5, head_and_neck, face, chest, abdomen_and_pelvic_contents, extremities, external)
## The function below takes as input one ais column and splits each
## row into a vector (this is what strsplit does). The split happens
## where there is a comma. It then calculates the number of elements
## in that vector that are equal to or greater than 3.
get_number_serious_injuries <- function(body.region.column) {
    sapply(strsplit(body.region.column, ","), function(x) sum(x >= 3))
}
ais.nsi <- do.call(cbind, lapply(ais, get_number_serious_injuries))
total.nsi <- rowSums(ais.nsi, na.rm = TRUE)
adtbi5["noofseriousinjury"] <- ais["noofseriousinjuries"] <-  total.nsi
nsi.kts <- ifelse(total.nsi == 0, 3,
           ifelse(total.nsi == 1 , 2,
           ifelse(total.nsi >= 2, 1, NA)))
## Calculate KTS
kts <- adtbi5["kts"] <- age.kts + sbp.kts + rr.kts + avpu.kts + nsi.kts

## creating a column for isolated tbi = 0 and tbi with polytrauma >0
ais1 <-select(adtbi5, pid,head_and_neck, face, chest, abdomen_and_pelvic_contents, extremities, external)
# separating contents of a column into different columns
ais1<- ais1 %>% separate(head_and_neck, c("a1"))
ais1<- ais1 %>% separate(face, c("b1"))
ais1<- ais1 %>% separate(chest, c("c1"))
ais1<- ais1 %>% separate(abdomen_and_pelvic_contents, c("d1"))
ais1<- ais1 %>% separate(extremities, c("e1"))
ais1<- ais1 %>% separate(external, c("f1","f2"))
ais1 <- ais1 %>% mutate_all(funs(replace_na(.,0)))
#View(ais1)
## polytrauma >= 1 and isolated tbi = 0 
poly1 <- ais1["p1"] <- with(ais1, ifelse(a1 >= 2 & b1 >= 1,1,0))
poly2<- ais1["p2"]<- with(ais1, ifelse(a1 >= 2 & c1 >= 1,1,0))
poly3 <- ais1["p3"] <- with(ais1,ifelse(a1 >= 2 & d1 >= 1,1,0))
poly4 <- ais1["p4"] <- with(ais1,ifelse(a1>= 2 & e1 >= 1, 1,0 ))                      
poly5 <-adtbi5["polytraumaorislotedtbi"] <-  ais1["polytraumaorislotedtbi"] <- rowSums(ais1[,c("p1","p2","p3","p4")], na.rm = T)

poly6 <- ais1["polytraumaorislotedtbi"] <-ifelse(ais1$a1 >= 1 & ais1$polytraumaorislotedtbi == 0 , "TBI",
                               ifelse(ais1$a1 >= 1 & ais1$polytraumaorislotedtbi >= 1 , "Polytrauma", NA))

adtbi5["TBI"] <- poly6


## performing complete case analysis to remove NA from the complete dataset of selected variables needed for analysis
tbi_241 <- adtbi5
tbi_242 <- tbi_241[!is.na(tbi_241$died),]
tbi_243 <- tbi_242[!is.na(tbi_242$gcs_t_1),]
tbi_244 <- tbi_243[!is.na(tbi_243$mgap),]
tbi_245 <- tbi_244[!is.na(tbi_244$gap),]
tbi_246 <- tbi_245[!is.na(tbi_245$rts),]
tbi_247 <- tbi_246[!is.na(tbi_246$kts),]
tbi_248 <- tbi_247[!is.na(tbi_247$TBI),]
adtbi <- tbi_248[!is.na(tbi_248$tran),]

nrow(adtbi)

patient.excluded <- nrow(adtbi1)-nrow(adtbi)

## Find no and % of males and females
m <- adtbi$sex == "Male"
n.male <- sum(m)
p.male <- round(mean(m) * 100)
f <- adtbi$sex == "Female"
n.f <- sum(f)
## Summarize age 
ages  <- adtbi$age
ages <- as.numeric(ages)
m.ages <- median(ages)
q1.ages <- quantile(ages,0.25)
q3.ages <- quantile (ages,0.75)
range.ages <- range(ages)
## Break age into groups
adtbi["agegroup"] <- agegroup <- cut(ages,
                                     breaks = c(17, 25, 45, 65, 85),
                                     labels =  c("18-24", "25-44", "45-64", "65+"))
table1 <- table(agegroup)
## Find percentage transferred
t <- adtbi$tran == "Yes"
n.t <- sum(t,na.rm=T)
p.t <- round(mean(t,na.rm=T)*100)
## Find mode of transport 
am <- adtbi$mot == "Ambulance"
n.am <- sum(am,na.rm=T)
p.am <- round(mean(am,na.rm=T)*100)
## Find percentage of Type of injury
ti <- adtbi$ti == "Blunt"
n.ti <- sum(ti,na.rm=T)
p.ti <- round(mean(ti,na.rm=T)*100)
## Create GCS group
a.gcs <- adtbi$gcs_t_1
## Break GCS into groups
adtbi["gcsgroup"] <- gcsgroup <- cut(a.gcs,
                                     breaks = c(2, 8, 13, 15),
                                     labels = c( "3-8", "9-13", "14-15"))

## Find number of surgery done
tos <- adtbi$tos 
surg1<- ifelse(tos == 0, 0,
              ifelse(tos == 999, 0, 1))
surgery1 <- adtbi["Surgery"] <- surg1 == 1
n.s <- sum(surgery1, na.rm = TRUE)
p.s <- round(mean(surgery1, na.rm = TRUE)*100)
## Find number of RTA and percentage of RTA
moi <- adtbi$moi
moi[grep("Road traffic injury", moi)] <- "Road traffic injury"
adtbi$moi <- moi
rta <- moi == "Road traffic injury"
n.rta <- sum(rta, na.rm = TRUE)
p.rta <- round(mean(rta, na.rm = TRUE) * 100)
## Percentage of fall
fall <- adtbi$moi == "Fall"
n.fall <- sum(fall,na.rm = TRUE)
p.fall <- round(mean(fall, na.rm = TRUE) * 100)
## Find mortality
died <- adtbi$died == "Yes"
n.died <- sum(died, na.rm = TRUE)
p.died <- round(mean(died, na.rm = TRUE) * 100)

## to find mortality at 24 hours
s24h <- adtbi$s24h == "Yes"
n.s24h <- sum(s24h, na.rm = T)
p.s24h <- round(mean(s24h, na.rm = TRUE) * 100)

## to find mortality at 30 days
s30d <- adtbi$s30d == "Yes"
n.s30d <- sum(s30d, na.rm = T)
p.s30d <- round(mean(s30d, na.rm = TRUE) * 100)

## Find discharged against medical advice
dama <- adtbi$dama == "Yes"
n.dama <- sum(dama,na.rm = T)
p.dama <- round(mean(dama,na.rm = T)*100)

## median & IQR
# GCS

gcs <- adtbi$gcs_t_1
m.gcs <- median(gcs)
q1.gcs <- quantile(gcs,0.25)
q3.gcs <- quantile (gcs,0.75)
range.gcs <- range(gcs)


# MGAP
mgap1 <- adtbi$mgap
m.mgap <- median(mgap1)
q1.mgap <- quantile(mgap1,0.25)
q3.mgap <- quantile (mgap1,0.75)
range.mgap <- range(mgap1)

# GAP
gap1 <- adtbi$gap
m.gap <- median(gap1)
q1.gap <- quantile(gap1,0.25)
q3.gap <- quantile (gap1,0.75)
range.gap <- range(gap1)

# RTS
rts1 <- adtbi$rts
m.rts <- median(rts1)
q1.rts <- quantile(rts1,0.25)
q3.rts <- quantile (rts1,0.75)
range.rts <- range(rts1)

#kts
kts1 <- adtbi$kts
m.kts <- median(kts1)
q1.kts <- quantile(kts1,0.25)
q3.kts <- quantile (kts1,0.75)
range.kts <- range(kts1)





## Creating Table for TBI patients demography using table one
table1.data <- with(adtbi,
                    data.frame("Age" = age,
                               "Age group" = agegroup,
                               "Gender" = sex,
                               "Mechanism of injury" = moi,
                               "Type of injury" = ti,
                              "Transferred" = ifelse(tran == "Yes", "Transferred", "Direct"),
                              "Systolic blood pressure" = sbp_1,
                              "Heart rate" = hr_1,
                              "Saturation" = spo2_1,
                              "Respiratory rate" = rr_1,
                              "Glasgow coma scale" = gcsgroup,
                              "Surgery" = ifelse(tos == "0", "Conservative", "Operative"),
                             "Mortality" = ifelse(died == "Yes", "Died", "Survived"),
                              check.names = FALSE
                              ))

x1 <- c("Age","Systolic blood pressure","Heart rate","Saturation","Respiratory rate")
a.table3 <- CreateTableOne(data=table1.data)
a.table4 <- knitr::kable(print(a.table3, nonnormal = x1, caption = "Table 1. Characteristics of TBI Patients", showAllLevels = TRUE, printToggle = FALSE))



## Analysis
## Length of hospital stay
doar.toar <- as_datetime(paste(adtbi$doar, adtbi$toar), format = "%Y-%m-%d %H:%M")
dodd.todd <- as_datetime(paste(adtbi$dodd, adtbi$todd), format = "%Y-%m-%d %H:%M")
loh <- adtbi["loh"] <- as.numeric(difftime(dodd.todd, doar.toar, units = "hours"))
summary(loh)
m.loh <- median(loh)
q1.loh <- quantile(loh,0.25, na.rm = TRUE)
q3.loh <- quantile(loh,0.75, na.rm = TRUE)
range.loh <- range(loh, na.rm = TRUE)

## Create status at 24 hours
died <- adtbi$died
## adtbi <- adtbi %>% filter_at(vars(65),any_vars(!is.na(.)))
ad.s24h <- adtbi["s24h"]<- ifelse(adtbi$loh  <= 24 & adtbi$died == "Yes", "Yes", "No")
dd <- adtbi$s24h == "Yes"
n.dd <- sum(dd, na.rm = TRUE)
p.dd <- round(mean(dd,na.rm = TRUE) * 100)
## Create status at 30 days 
ad.s30d <- adtbi["s30d"] <- ifelse(adtbi$loh <= 24*30 & adtbi$died == "Yes", "Yes", "No")
dd3 <- adtbi$s30d == "Yes"
n.dd3 <- sum(dd3, na.rm = T)
p.dd3 <- round(mean(dd3,na.rm=T)*100)

##creating 24hrs in hospital mortality dataset 
tbi_24 <- select(adtbi,s24h, gcs_t_1, mgap,gap,rts,kts, tran, ti,TBI)


## creating 30days in-hospital mortality dataset
 tbi_30 <- select(adtbi,s30d,gcs_t_1, mgap,gap,rts,kts)

## Creating ROC curve  for in hospital mortality


### Added two functions, ROC.func and ROC.score for Plots and AUC calculations:
ROC.func <- function(x, y, z) {
  input <- select(z,x,y)
  input <- input[complete.cases(input),]   
  input[, x, drop = TRUE] <- as.factor(input[, x, drop = TRUE])
  input[, y, drop = TRUE] <- as.factor(input[, y, drop = TRUE])
  GLM <- glm(input[, x, drop = TRUE] ~ input[, y, drop = TRUE], family = binomial)
  par(pty= "s")
  return(roc(input[, x, drop = TRUE],GLM$fitted.values, plot = TRUE, legacy.axes= TRUE))
}

## ROC analysis for 30 days mortality

## ROC Curve for gcs

rocgcs30 <- ROC.func("s30d","gcs_t_1",tbi_30)
ci.auc(rocgcs30)
## ROC Curve for mgap
rocmgap30 <- ROC.func("s30d","mgap", tbi_30)
ci.auc(rocmgap30)
## ROC curve GAP
rocgap30 <- ROC.func("s30d","gap", tbi_30)
ci.auc(rocgap30)
## ROC curve RTS s24hrs
rocrts30 <- ROC.func("s30d","rts", tbi_30)
ci.auc(rocrts30)
## Roc KTS
rockts30 <- ROC.func("s30d","kts", tbi_30)
ci.auc(rockts30)
## creating a combine plot for all 5 roc curve s30d
par(pty= "s")
plot(rocgcs30, col = 3, lty = 5, main = "ROC", legacy.axes= TRUE)## green
plot(rocmgap30, col = 4, lty = 3, add = TRUE) ## blue
plot(rocgap30, col = 6, lty = 4, add = TRUE) ## pink
plot(rocrts30, col = 7, lty = 5, add = TRUE) ## yellow
plot(rockts30,col = 9 , lty= 9, add = TRUE)
## P value for different ROC curve
# comparing ROC curve among s30d
r30days1 <- roc.test(rocgcs30,rocmgap30)
r30days2 <- roc.test(rocgcs30,rocgap30)
r30days3 <- roc.test(rocgcs30,rocrts30)
r30days4<- roc.test(rocgcs30,rockts30)



## ALL ROC curve analysis done below is for 24 hrs mortality

## ROC Curve for gcs

rocgcs3 <- ROC.func("s24h","gcs_t_1",tbi_24)
ci.auc(rocgcs3)
## ROC Curve for mgap
rocmgap <- ROC.func("s24h","mgap", tbi_24)
ci.auc(rocmgap)
## ROC curve GAP
rocgap <- ROC.func("s24h","gap", tbi_24)
ci.auc(rocgap)
## ROC curve RTS s24hrs
rocrts <- ROC.func("s24h","rts", tbi_24)
ci.auc(rocrts)
## Roc KTS
rockts24 <- ROC.func("s24h","kts", tbi_24)
ci.auc(rockts24)
## creating a combine plot for all 5 roc curve
par(pty= "s")
plot(rocgcs3, col = 3, lty = 5, main = "ROC", legacy.axes= TRUE)## green
plot(rocmgap, col = 4, lty = 3, add = TRUE) ## blue
plot(rocgap, col = 6, lty = 4, add = TRUE) ## pink
plot(rocrts, col = 7, lty = 5, add = TRUE) ## yellow
plot(rockts24,col = 9 , lty= 9, add = TRUE)
## P value for different ROC curve
# comparing ROC curve among s24h
r1 <- roc.test(rocgcs3,rocmgap)
r2 <- roc.test(rocgcs3,rocgap)
r3 <- roc.test(rocgcs3,rocrts)
rkts1 <- roc.test(rocgcs3,rockts24)

## Subgroup analysis as suggested on thursday presentation
## subgroup analysis - transfer status

# creating a dataset with directly admitted patients
addirect <- filter(tbi_24, tran == "No")
nrow(addirect)


# creating a dataset with transferred patients
adtrans <- filter(tbi_24, tran == "Yes")
nrow(adtrans)


# roc curve gcs direct patients for 24hrs 


roc9 <- ROC.func("s24h","gcs_t_1", addirect)
ci.auc(roc9)
# roc curve mgap direct patients

roc10 <- ROC.func("s24h","mgap", addirect)
ci.auc(roc10)
# roc curve gap direct patients

roc11 <- ROC.func("s24h","gap", addirect)
ci.auc(roc11)
# roc curve rts direct patients

roc12 <- ROC.func("s24h","rts", addirect)
ci.auc(roc12)
## roc for KTS and directly admitted patients

rocktsd <- ROC.func("s24h","kts", addirect)
ci.auc(rocktsd)

## creating a combine plot for all 4 roc curve for directly admitted patients
par(pty= "s")
plot(roc9, col = 3, lty = 5, main = "Direct admission", legacy.axes= TRUE)## green
plot(roc10, col = 4, lty = 3, add = TRUE) ## blue
plot(roc11, col = 6, lty = 4, add = TRUE) ## pink
plot(roc12, col = 7, lty = 5, add = TRUE) ## yellow
plot(rocktsd, col = 9, lty = 9, add = TRUE)

## plotting roc curve and calculating auc for transferred patients


roc13 <- ROC.func("s24h","gcs_t_1", adtrans)
ci.auc(roc13)
# roc curve mgap transferred patients

roc14 <- ROC.func("s24h","mgap", adtrans)
ci.auc(roc14)
# roc curve gap transferred patients

roc15 <- ROC.func("s24h","gap", adtrans)
ci.auc(roc15)
# roc curve rts

roc16 <- ROC.func("s24h","rts", adtrans)
ci.auc(roc16)
# roc kts and transferred

rocktst <- ROC.func("s24h","kts", adtrans)
ci.auc(rocktst)
## creating a combine plot for all 4 roc curve
par(pty= "s")
plot(roc13, col = 3, lty = 5, main = "Transferred admission", legacy.axes= TRUE)## green
plot(roc14, col = 4, lty = 3, add = TRUE) ## blue
plot(roc15, col = 6, lty = 4, add = TRUE) ## pink
plot(roc16, col = 7, lty = 5, add = TRUE) ## yellow
plot(rocktst, col = 9, lty = 9, add = TRUE)

# comparing ROC curve for directly admitted patients
r7 <- roc.test(roc9,roc10)
r8 <- roc.test(roc9,roc11)
r9 <- roc.test(roc9,roc12)
rktsd <- roc.test(roc9,rocktsd)
# comparing ROC curve for transferred patients
r10 <- roc.test(roc13,roc14)
r11 <- roc.test(roc13,roc15)
r12 <- roc.test(roc13,roc16)
rktst <- roc.test(roc13, rocktst)


## subgroup polytrauma and isolated TBI

## creating dataset for polytrauma patients
polytrauma <- filter(tbi_24, TBI == "Polytrauma" )
nrow(polytrauma)

## plotting roc curve and calculating auc for polytrauma patients patients

roc30 <- ROC.func("s24h","gcs_t_1", polytrauma)
ci.auc(roc30)
## roc curve mgap polytrauma

roc31 <- ROC.func("s24h","mgap", polytrauma)
ci.auc(roc31)
# roc curve gap polytrauma patients

roc32 <- ROC.func("s24h","gap", polytrauma)
ci.auc(roc32)
# roc curve rts polytrauma patients

roc33 <- ROC.func("s24h","rts", polytrauma)
ci.auc(roc33)
# roc kts and polytrauma patients

roc34 <- ROC.func("s24h","kts", polytrauma)
ci.auc(roc34)

## creating a combine plot for all 4 roc curve for polytrauma patients
par(pty= "s")
plot(roc30, col = 3, lty = 5, main = "Ploytrauma (Including TBI)", legacy.axes= TRUE)## green
plot(roc31, col = 4, lty = 3, add = TRUE) ## blue
plot(roc32, col = 6, lty = 4, add = TRUE) ## pink
plot(roc33, col = 7, lty = 5, add = TRUE) ## yellow
plot(roc34, col = 9, lty = 9, add = TRUE)

# comparing ROC curve for polytrauma patients
r19 <- roc.test(roc30,roc31)
r20 <- roc.test(roc30,roc32)
r21 <- roc.test(roc30,roc33)
r22 <- roc.test(roc30, roc34)

## creating dataset for isolated TBI
isolated <- filter(tbi_24, TBI == "TBI" )
nrow(isolated)

## plotting roc curve and calculating auc for isolated TBI patients

roc35 <- ROC.func("s24h","gcs_t_1", isolated)
ci.auc(roc35)
## roc curve mgap polytrauma

roc36 <- ROC.func("s24h","mgap", isolated)
ci.auc(roc36)
# roc curve gap polytrauma patients

roc37 <- ROC.func("s24h","gap", isolated)
ci.auc(roc37)
# roc curve rts polytrauma patients

roc38 <- ROC.func("s24h","rts", isolated)
ci.auc(roc38)
# roc kts and polytrauma patients

roc39 <- ROC.func("s24h","kts", isolated)
ci.auc(roc39)

## creating a combine plot for all 4 roc curve for isolated TBI patients
par(pty= "s")
plot(roc35, col = 3, lty = 5, main = "Isolated TBI", legacy.axes= TRUE)## green
plot(roc36, col = 4, lty = 3, add = TRUE) ## blue
plot(roc37, col = 6, lty = 4, add = TRUE) ## pink
plot(roc38, col = 7, lty = 5, add = TRUE) ## yellow
plot(roc39, col = 9, lty = 9, add = TRUE)

# comparing ROC curve for polytrauma patients
r23 <- roc.test(roc35,roc36)
r24 <- roc.test(roc35,roc37)
r25 <- roc.test(roc35,roc38)
r26 <- roc.test(roc35, roc39)





 
 ## creating flow chart of sample size
 
Flow.chart <-  grViz("digraph {
      graph[layout = dot, rankdir = RR]
      node [fontname = Flowchart, shape = rectangle]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']

      # edge definitions with the node IDs
      tab1 -> tab2;
      tab2 -> tab3;
      tab2 -> tab4;
      tab4 -> tab5;
      tab4 -> tab6
      }

      [1]: 'TITCO Dataset n= 16000'
      [2]: 'All patients with Traumatic brain injury n= 9268 '
      [3]: 'patients with age < 18 was excluded n= 2024'
      [4]: 'Patients with age greater than or equal to 18 was included n= 7244'
      [5]: 'Incomplete Data were excluded n= 3938'
      [6]: 'Study sample n=3306'
      ")
 
 
 

 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 

 
```

##Results
A total of `r nrow(adtbi)` TBI patients were included for our study. `r n.male` (`r p.male`%) were male. Median age of patients was `r m.ages` (quantile range: `r q1.ages` to `r q3.ages` & range: `r range.ages`).   Majority of the patients were within the GCS range 14-15 (Table 1). Surgery was performed in `r n.s` (`r p.s`%) patients. Road traffic injuries (RTA) accounted for the majority of TBI `r n.rta`(`r p.rta`%) followed by falls `r n.fall`(`r p.fall`%). Mortality was `r n.died`(`r p.died`%) among these patients.GCS discriminated significantly better than KTS, there were no significant differences in discrimination comparing GCS with MGAP, RTS but GAP score discriminated better than GCS.  
## Conclusion
GCS was comparable to other scores in discrimination of in-hospital mortality.

##Introduction
Traumatic Brain Injury (TBI) is defined as “A disruption or alteration of brain function due external forces”.(1) Globally, there was a 3.6% increase in age standardised incidence rates among TBI patients when compared between 1990 and 2016.(2) In India, approximately 10 million people suffer from moderate to severe TBI each year.(3) Rapid urbanization and an increase in the number of motor vehicles are important causes of the increase in incidence of TBI among the adult population in India.(4) Road traffic injuries are the commonest mechanism of injury among these patients.(4) 
There are different types of trauma scoring systems, keeping anatomical, physiological and combined systems into consideration. (5) They are used widely for describing injury severity, prognosis of the patients, triage, research and registry. (6) GCS is the most commonly used scoring system used in TBI and is a good predictor of outcome among TBI patients. (7,10,11,12) GCS is easy to use by members of health care team (8) and provides a mechanism to quickly evaluate the severity of TBI.(9) GCS can also be used to see changes in the level of consciousness at different period of time.(8) 
GCS alone is a very robust scoring system. But it has been included as an integral component in some other trauma scoring systems like MGAP and GAP. The acronym MGAP stands for Mechanism (of injury),GCS, Age, (systolic blood) Pressure. GAP is different from MGAP, in that it excludes mechanism of injury. Whether these newer scores outperform GCS alone is unclear.Thus the purpose of this study is to compare GCS alone with other trauma scoring systems viz. MGAP, GAP, Revised Trauma Score (RTS) and Kampala Trauma Score (KTS)in adult TBI patients.


##Method:

#Setting:
The study was an analysis of prospectively collected data from the “Towards improvement of trauma care outcomes in India” (TITCO) Cohort. This data was collected between2013 to 2015 at four urban tertiary care referralcentres spread across India. Of the 4 centres,  All India Institute of Medical Science (AIIMS), New Delhi, King Edward Memorial Hospital and Seth Gordhandas Sunderdas Medical College (KEM), Mumbai, Lokmanya Tilak Municipal General Hospital and Medical College (Sion hospital) had a dedicated trauma care facility, while  Seth Sukhlal Karnani Memorial Hospital/Institute of Post Graduate Medical Education & Research, Kolkata, though a tertiary referral centre, didn’t have  a dedicated trauma care facility during the study period
One project officer per hospital was appointed to collect data prospectively in 3 eight hour shifts in the emergency room. In-order to minimise bias and maintain consistency in data collection, weekly telephonic meetings were held where problems were discussed and regular quality control checks done.

#Inclusion criteria:
Patients with age of 18 years and above with the history of TBI admitted to the hospital.
#Exclusion criteria
Patient with incomplete data and who were lost to follow-up was removed from the study.
#Variables:
Inthis study five trauma scoring systems were compared - GCS, MGAP, GAP, RTS and KTS. All these scores have been published in English literatureand validated (13, 14, 15, 16). All values were calculated as per published formulae. Vitals such as systolic blood pressure, respiratory rate, GCS were measured on arrival atthe hospital. Demographic variables such as age, gender, mechanism of injury, type of injurywas collected from the patients’ records. For calculation of KTS,AVPU score was calculated by extrapolating the GCS score (GCS 14–15 = ‘‘alert’’, GCS 10–13 = ‘‘responds to voice’’, GCS 5–9 = ‘‘responds to pain’’, GCS 3–4 = ‘‘unresponsive’’). (17) Number of serious injuries wascalculated from the patient’s records.
#Outcome:
Primary outcome of this study is 24 hours in-hospital mortality. Secondary outcome of this study is 30 days in- hospital mortality.
#StatisticalAnalysis:
Data was analysed using R version 4.0.2(18). Median (IQR) was used to describe quantitative variables, while count and percentage were used for qualitative variables.If any data for any of the 5 scores was missing, that patient was excluded from the analysis. Area under the receiver operating characteristics (ROC) curve was used to compare the performance of the different trauma scoring system. A p-value of less than 0.05 was determined to be statistically significant.




#Results
A total of `r nrow(titco)` patients were included during the study period. Of these, `r nrow(titcotbi)` patients were admitted with traumatic brain injury (TBI). From this, `r nrow(adtbi1)` patients were above the age of 17 and was eligible for our study.Furthermore, we had to exclude `r patient.excluded` patients as they had missing data for systolic blood pressure, respiratory rate,GCS, mechanism of injury. The final sample size for the study was `r nrow(adtbi)`.  (Fig 1) 

```{r Flow.chart, fig.align = 'center', out.width = "100%", fig.cap = "Flow chart showing inclusion and exclusion criteria and sample size", echo = FALSE}

```


The median age was `r m.ages` (quantile range: `r q1.ages` to `r q3.ages` & range: `r range.ages`) and the majority of them were male `r n.male` (`r p.male`%). Most of the patients were admitted with road traffic injuries `r n.rta`(`r p.rta`%) followed by falls `r n.fall`(`r p.fall`%). A total of `r n.t`(`r p.t`%) patients were transferred from other hospitals and `r n.s` (`r p.s`%) patients had operative intervention.In hospital 24 hours and 30 days mortality was `r n.s24h`(`r p.s24h`) and  `r n.s30d`(`r p.s30d`) respectively. Out of all patients with TBI `r n.died`(`r p.died`%) died. (Table1) The median, interquartile range (IQR), range of GCS, MGAP, GAP, RTS and KTS are shown in Table 2.

```{r a.table4, echo = FALSE}
 kableone(a.table4,caption = "Characteristics of patients with Traumatic brain injury included in this study", label = "table1")
```

ROC analysis was performed to compare the discrimination of GCS, MGAP, GAP, RTS and KTS with regards to 24 hrs in-hospital mortality and 30 days in-hospital mortality.In-case of 24 hours mortality, GCS discriminated significantly better compared with KTS (p= `r rkts1[["p.value"]]`) but there were no significant differences in the discrimination of GCS when compared with MGAP (p= `r r1[["p.value"]]`) and rts (p= `r r3[["p.value"]]`). GAP discriminated better when compared with GCS (p= `r r2[["p.value"]]`). For 30 days mortality, GCS discriminated significantly better compared with KTS (p= `r r30days4[["p.value"]]`) but there were no significant differences in the discrimination of GCS when compared with MGAP (p= `r r30days1[["p.value"]]`) and GAP (p= `r r30days2[["p.value"]]`). RTS discriminated better when compared with GCS (p= `r r30days4[["p.value"]]`). (Table 2, Fig 2)

In 24 hours mortality, we also compared isolated TBI with polytrauma (including TBI) and transferred patients with directly admitted patients.In both cases, GCS discriminated significantly better than KTS. No significant difference in the discrimination was seen when compared to MGAP, GAP, RTS scores. (Fig 3)

Table 2: 
|           |                                           |           24 hours Mortality                         |                 30 days Mortality                    |
|           |                                           |--------------------------|---------------------------|------------------------------------------------------|
|           |Median (IQR)                               | 95% Confidence interval  | Area under the curve(AUC) | 95% Confidence interval  | Area under the curve(AUC) | 
|-----------|-------------------------------------------|--------------------------|---------------------------|--------------------------|---------------------------|
| GCS       | `r m.gcs`(`r q1.gcs`-`r q3.gcs`)          | `r ci.auc(rocgcs3)`      |  `r rocgcs3[["auc"]]`     | `r ci.auc(rocgcs30)`     |  `r rocgcs30[["auc"]]`     |
| MGAP      | `r m.mgap`(`r q1.mgap`-`r q3.mgap`)       | `r ci.auc(rocmgap)`      |  `r rocmgap[["auc"]]`     | `r ci.auc(rocmgap30)`    |  `r rocmgap30[["auc"]]`    |
| GAP       | `r m.gap`(`r q1.gap`-`r q3.gap`)          | `r ci.auc(rocgap)`       |  `r rocgap[["auc"]]`      | `r ci.auc(rocgap30)`     |  `r rocgap30[["auc"]]`     |
| RTS       | `r m.rts`(`r q1.rts`-`r q3.rts`)          | `r ci.auc(rocrts)`       |  `r rocrts[["auc"]]`      | `r ci.auc(rocrts30)`     |  `r rocrts30[["auc"]]`     |
| KTS       | `r m.kts`(`r q1.kts`-`r q3.kts`)          | `r ci.auc(rockts24)`     |  `r rockts24[["auc"]]`    | `r ci.auc(rockts30)`     |  `r rockts30[["auc"]]`     |




Table 3:


|           |           Isolated TBI                               |                 Poly-trauma (including TBI)          |
|           |--------------------------|---------------------------|------------------------------------------------------|
|           | 95% Confidence interval  | Area under the curve(AUC) | 95% Confidence interval  | Area under the curve(AUC) | 
|-----------|--------------------------|---------------------------|--------------------------|---------------------------|
| GCS       | `r ci.auc(roc35)`        |  `r rocgcs3[["auc"]]`     | `r ci.auc(roc30)`        |  `r rocgcs30[["auc"]]`    |
| MGAP      | `r ci.auc(roc36)`        |  `r rocmgap[["auc"]]`     | `r ci.auc(roc31)`        |  `r rocmgap30[["auc"]]`   |
| GAP       | `r ci.auc(roc37)`        |  `r rocgap[["auc"]]`      | `r ci.auc(roc32)`        |  `r rocgap30[["auc"]]`    |
| RTS       | `r ci.auc(roc38)`        |  `r rocrts[["auc"]]`      | `r ci.auc(roc33)`        |  `r rocrts30[["auc"]]`    |
| KTS       | `r ci.auc(roc39)`        |  `r rockts24[["auc"]]`    | `r ci.auc(roc34)`        |  `r rockts30[["auc"]]`    |



Table 4:

|           |           Directly admitted Patients                 |                 Transferred Patients                 |
|           |--------------------------|---------------------------|------------------------------------------------------|
|           | 95% Confidence interval  | Area under the curve(AUC) | 95% Confidence interval  | Area under the curve(AUC) | 
|-----------|--------------------------|---------------------------|--------------------------|---------------------------|
| GCS       | `r ci.auc(roc9)`         |  `r rocgcs3[["auc"]]`     | `r ci.auc(rocgcs13)`     |  `r rocgcs30[["auc"]]`    |
| MGAP      | `r ci.auc(roc10)`        |  `r rocmgap[["auc"]]`     | `r ci.auc(rocmgap14)`    |  `r rocmgap30[["auc"]]`   |
| GAP       | `r ci.auc(roc11)`        |  `r rocgap[["auc"]]`      | `r ci.auc(rocgap15)`     |  `r rocgap30[["auc"]]`    |
| RTS       | `r ci.auc(roc12)`        |  `r rocrts[["auc"]]`      | `r ci.auc(rocrts16)`     |  `r rocrts30[["auc"]]`    |
| KTS       | `r ci.auc(rocktsd)`      |  `r rockts24[["auc"]]`    | `r ci.auc(rocktst)`      |  `r rockts30[["auc"]]`    |













```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
