---
title: Profile of patients triaged green in the Emergency Department (ED) of a secondary care hospital in Mumbai
author: "Anna"
date: "12/05/2020"
output:
  word_document: default
  html_document: 
    fig_caption: yes
    fig_height: 3
    fig_width: 5
---


```{r setup, include=FALSE, results='asis', message=FALSE}
library(knitr)
library(dplyr)
library(psych)
library(data.table) 
library(tableone)
library(epiDisplay)
library(tigerstats)
library(splitstackshape)
library(tidyverse)
library(tidyr)
opts_chunk$set(echo = FALSE)
```

# Abstract

## Introduction

Injuries are one of the leading causes of emergency department visits.  Based on injury severity and physiological parameters patients are triaged into red, orange, yellow, and green, from most to least urgent. Emergency departments provide quick and easy access to healthcare, leading to visits of patients with no injury or minor injuries. Although these patients are triaged green they consume significant resources. It is important to understand the characteristics of these patients to understand how they should be managed so that resources can be directed towards serious patients.

## Aim

The aim of this study is to describe the demographic, physiological and injury characteristics of trauma patients triaged green in the ED of a secondary healthcare hospital in Mumbai.

## Methods

Data of patients triaged green were extracted from the Trauma Triage Study in India dataset. A retrospective cohort analysis of the data was performed using R. 

## Results
```{r Data_cleaning_manipulation, warning=FALSE, results='hide'}

# Get data directory

data.dir <- getOption("data.dir")
if (is.null(data.dir))
    stop ("Please set the directory where the data is kept using options(data.dir = \"<path to data directory>\")")

## datatable with ISS
g <- data.table::fread(file.path(data.dir, "ttris-dataset-with-iss-8317-20190708120045.csv"))

## datatable without ISS
h <- data.table::fread(file.path(data.dir, "TTRIS.csv"))

## Merged the two datatables
h$pid <- as.integer(h$pid)

merged <- merge(g, h, by = "pid")

## Duplicated rows
merged <- merged[!duplicated(merged$pid), ]

## Changed column names
names(merged)[names(merged) == "iss.y"] <- "delcolumn.x"
names(merged)[names(merged) == "iss.x"] <- "iss.y"

## Removing duplicated columns
dt <- data.table(merged)
drop.cols <- grep(".x$", colnames(dt))
dt[, (drop.cols) := NULL]

## Removed .y from the coulmn names
names(dt) <- sub('\\.[^.]+$', "", names(dt))

## Rearranging columns
dt <- dt %>% dplyr::select(centre,pid,cr,qcrec:vsqs,ic:toa,egcs:rr,nsi,tc,dvsr:tvsr,dus:testvar,locked,X0:iss,doi_toi:valid)

extra_col <- dt %>% dplyr::select(doi_toi:valid)

## Adding the extra columns to g
g <- cbind(g,extra_col)

## Changing class of the columns
g <- as.data.frame(g)
dt <- as.data.frame(dt)

dt[] <- lapply(names(dt), function(name) {
    column <- dt[, name]
    correct.class <- class(g[, name])
    if (correct.class == "factor") {
        new.column <- factor(column, levels = levels(g[, name]))
    } else {
        new.column <- as(column, correct.class)
    }
    return (new.column)
})

## dplyr::all_equal(g,dt) ##To compare the two dataframes

# Creating age groups
agebreaks <- c(17,25,45,65,98)
agelabels <- c("18-24","25-44","45-64","65+")
setDT(dt)[,agegrp := cut(dt$age,
                           breaks = agebreaks,
                           right = FALSE,
                           labels = agelabels)]

# Mechanism of Injury
moi.collapsed <- moi <- as.character(dt$moi)

get_icd_range <- function(first.char, start.number, end.number) {
    icd.range <- paste0(
        paste0(
            "^",
            first.char,
            stringr::str_pad(start.number:end.number, 2, pad = "0"),
            "[0-9]?$"),
        collapse = "|")
    return (icd.range)
}

icd.ranges <- list(c("Transport accident" = get_icd_range("V", 0, 99)),
                   c("Fall" = get_icd_range("W", 0, 19)),
                   c("Animal bite" = get_icd_range("W", 50, 64)),
                   c("Assault" = paste0(get_icd_range("X", 85, 99), "|", get_icd_range("Y", 0, 9))))

for(icd.range in icd.ranges) moi.collapsed[grep(icd.range, moi)] <- names(icd.range)
moi.collapsed[!(moi.collapsed %in% sapply(icd.ranges, attr, "names"))] <- "Other"

dt$moi.collapsed <- as.factor(moi.collapsed)

# Calculating ISS
dt <- as.tibble(dt)
ais <- dt %>% dplyr:: select(starts_with("clean_"))

ais.list <- lapply(names(ais), function(x) cSplit(ais[, x], x, sep = ","))
row_max <- function(x, na.rm = FALSE) {
    if (all(is.na(x)))
        return (NA)
    max(x, na.rm = na.rm)
}
ais.max <- lapply(ais.list, function(x) apply(x, 1, row_max, na.rm = TRUE))
get_top_3 <- function(x) {
    x <- sort(x, decreasing = TRUE, na.last = TRUE)
    i <- min(c(3, length(x)))
    x[seq_len(i)]
}
calculate_iss <- function(x) {
    top_3 <- get_top_3(x)
    iss <- sum(x[1]^2, x[2]^2, x[3]^2, na.rm = TRUE)
    top3 <- get_top_3(x)
    iss <- sum(top3[1]^2, top3[2]^2, top3[3]^2, na.rm = TRUE)
    if (all(is.na(top3)))
        iss <- NA
    iss
}
iss <- apply(do.call(cbind, ais.max), 1, calculate_iss)
dt$iss <- iss

## grouping ISS
isspv <- iss <- dt$iss

isspv[iss >= 16 & iss < 25] <- "Severe"
isspv[iss >= 9 & iss < 16] <- "Moderate"
isspv[iss > 0 & iss < 9] <- "Minor"
isspv[is.na(iss)] <- "No injuries defined"

dt$isspv <- isspv

# Append a new variable 'GCS' to the dataset
gcs <- with(dt, egcs + vgcs + mgcs)
gcs <- replace(gcs, gcs > 15, NA)

gcspv <- gcs
gcspv[gcs >= 13] <- 4
gcspv[gcs >= 9 & gcs < 13] <- 3
gcspv[gcs >= 6 & gcs < 9] <- 2
gcspv[gcs >= 4 & gcs < 6] <- 1
gcspv[gcs < 4] <- 0

## RR parameter value for calculating RTS
rr <- with(dt,replace(rr, rr == 999, NA))
rrpv <- rr

rrpv[rr >= 10 & rr <= 29] = 4
rrpv[rr > 29] = 3
rrpv[rr >= 6 & rr <= 9] = 2
rrpv[rr >= 1 & rr <= 5] = 1
rrpv[rr == 0] = 0

## SBP parameter value for calculating RTS
sbp <- with(dt,replace(sbp, sbp == 999, NA))
sbppv <- sbp

sbppv[sbp > 89] <- 4
sbppv[sbp >=76 & sbp <= 89] <- 3
sbppv[sbp >= 50 & sbp <= 75] <- 2
sbppv[sbp >= 1 & sbp <= 49] <- 1
sbppv[sbp == 0] <- 0

## Formula for calculating RTS
dt$RTS = round(0.9368*as.numeric(gcspv) + 
                 0.7326*as.numeric(sbppv) + 
                 0.2908*as.numeric(rrpv), digits = 0)

# Complete case analysis
dt$s24h <- replace(dt$s24h, dt$s24h == 999, 99) ##Loss of follow up 
dt$s30d <- replace(dt$s30d, dt$s30d == 999, 99)
dt$s6m <- replace(dt$s6m, dt$s6m == 999, 99)
dt$s <- replace(dt$s, dt$s == 999, 99)

## Converting all Injury 999 to Not Defined
inj_cols <- dt[,grep("inj", colnames(dt))]

inj_cols <- inj_cols %>% 
    mutate_all(funs(str_replace(., "999", "ND")))

dt[,grep("inj", colnames(dt))] <- inj_cols

## Converting NA to ND in max cols
#dt %>% dplyr:: select(starts_with("max_")) %>% 
#    mutate_all(funs(str_replace(., "NA", "ND")))

## Dropping unwanted columns
drop.cols2 <- dt %>% dplyr:: select(starts_with("clean_"), starts_with("max_"), qcrec, vsqs, testvar, locked, X0)

drop.cols2 <- colnames(drop.cols2)

dt <- dt[, !(colnames(dt) %in% c(drop.cols2))]

## Converting rest of the 999s to NA
dt[] <- dt[] %>% 
    mutate_all(funs(na_if(.,"999")))

missing.data.list <- lapply(dt, function(column) {
    n.na <- sum(is.na(column))
    p.na <- round((n.na/length(column)) * 100)
    missing <- data.frame("Count" = n.na, "Percentage" = p.na)
    return (missing)
})

missing.data <- do.call(rbind, missing.data.list)

maximum.missing <- rownames(missing.data)[missing.data$Count == max(missing.data$Count)]

complete.index <- complete.cases(dt)
n.incomplete <- sum(!complete.index)
p.incomplete <- round((n.incomplete/nrow(dt)) * 100)

dtt <- dt[complete.index, ]

```

```{r result2}

# Created a dataframe of triage category green
kbbh <- subset(x=dtt,subset = dtt$tc == 0)

# Number of records
dm <- as.numeric(nrow(kbbh))

# Age
kbbh$age <- as.numeric(kbbh$age)
kbbh$age <- replace(kbbh$age, kbbh$age == 999, NA)
mean_age <- round(mean(kbbh$age, na.rm = TRUE),2)
stddev_age <- round(sd(kbbh$age, na.rm = TRUE),2)

nm_age <- sum(kbbh$agegrp != "65+")
percent_agebelow65 <- round(mean(kbbh$agegrp != "65+") * 100, 2)

# Sex
nm_male <- sum(kbbh$sex == 1)
percent_male <- round(mean(kbbh$sex == 1) * 100,2)

nm_female <- sum(kbbh$sex == 0)
percent_female <- round(mean(kbbh$sex == 0) * 100,2)

# Type of Injury
nm_blunt <- sum(kbbh$tyi == 0)
percent_blunt <- round(mean(kbbh$tyi == 0) * 100,2)

# Mode of Transport
kbbh$mot <- replace(kbbh$mot, kbbh$mot == 2, "Private Vehicles")

nm_pvt <- sum(kbbh$mot == "Private Vehicles")
percent_mot_pvt <- round(mean(kbbh$mot == "Private Vehicles") * 100,2)

nm_onfoot <- sum(kbbh$mot == 3)
percent_mot_onfoot <- round(mean(kbbh$mot == 3) * 100,2)

sort_most_common <- function(v) {
   uniqv <- unique(v)
   uniqv[sort.list(tabulate(match(v, uniqv)), decreasing = TRUE)]
}

sorted_mot <- sort_most_common(kbbh$mot)
mot1 <- sorted_mot[[1]][]

# MOI
calculate_n_p_moi <- function(level.name) {
    moi.level <- dplyr::filter(kbbh, kbbh$moi.collapsed == level.name) # neumerator
    nm <- nrow(moi.level)
    percent <- round((nm/dm) * 100,2)
    return.list <- list(nm = nm, percent = percent)
    return (return.list)
}

moi.data <- lapply(setNames(nm = levels(kbbh$moi.collapsed)), calculate_n_p_moi)

p_animalbite <- moi.data[[1]][2]
p_assualt <- moi.data[[2]][2]
p_fall <- moi.data[[3]][2]
p_other <- moi.data[[4]][2]
p_tra <- moi.data[[5]][2]

sorted_moi <- sort_most_common(kbbh$moi.collapsed)

moi1 <- sorted_moi[[1]][]
moi2 <- sorted_moi[[2]][]
moi3 <- sorted_moi[[3]][]

# Trauma outcomes
kbbh$s6m <- as.factor(kbbh$s6m)

calculate_n_p_outcome <- function(level.name) {
    outcome.level <- dplyr::filter(kbbh, kbbh$s6m == level.name) # neumerator
    nm <- nrow(outcome.level)
    percent <- round((nm/dm) * 100,2)
    return.list <- list(nm = nm, percent = percent)
    return (return.list)
}

outcome.data <- lapply(setNames(nm = levels(kbbh$s6m)), calculate_n_p_outcome)

p_outcome_alive <- outcome.data[[1]][2]
p_outcome_dead <- outcome.data[[2]][2]
n_outcome_dead <- outcome.data[[2]][1]

p_followup <- as.numeric(p_outcome_alive)+as.numeric(p_outcome_dead) 

```

There was `r nrow(dt)` observations in the study data. Out of these, `r n.incomplete` (`r p.incomplete` %) observations were incomplete.

In the TTRIS dataset, `r dm` patients were triaged green. The mean age of these patients was `r mean_age` (+-`r stddev_age`). `r percent_agebelow65`% of the patients were below 65 years of age. `r percent_male`% patients were male, while `r percent_female`% patients were female. `r percent_blunt`% patients had blunt injuries. Most common mode of arrival to the ED was through `r mot1` (`r percent_mot_pvt`)%. `r percent_mot_onfoot`% patients came to the ED on foot. `r moi1` were the most common cause of injury `r p_assualt`% followed by `r moi2` `r p_tra`% and `r moi3` `r p_fall`%. Injury due to animal bites were. `r p_followup`% patients were successfully followed up at 6months, of which `r n_outcome_dead` patients were reported dead. 

```{r ISS_category, RTS}

Table1 <- colPerc(xtabs(~kbbh$isspv+kbbh$moi.collapsed, data = kbbh))
kable(Table2, 
      caption = "Table 1. Percentage distribution of Injury severity among different Mechanism of Injury",
)

p_iss_nd <- round(mean(kbbh$isspv == "No injuries defined") * 100,2)
p_iss_mi <- round(mean(kbbh$isspv == "Minor")*100,2)

Table2 <- rowPerc(xtabs(~kbbh$RTS, data = kbbh))
kable(Table2, 
      caption = "Table 1. RTS of patients",
)

RTS_8 <- dplyr::filter(kbbh, kbbh$RTS == 8) # neumerator
nm_RTS_8 <- nrow(RTS_8)
p_RTS_8 <- round((nm_RTS_8/ dm) * 100,2)

```

## Conclusion

Among the patients triaged green, injury due to assualt and transport accidents were predominant. `r p_RTS_8`% patients had an RTS score of 8. Patients with minor injuries and no injuries defined are `r p_iss_mi`% and `r p_iss_nd`% respectively. 

## Analysis

```{r Extra, warning=FALSE, results='hide'}

## Table of age group
tab1(kbbh$agegrp, main = "Distribution of patients by Age")
tab1(kbbh$sex, main = "Distribution of patients by Gender")
tab1(kbbh$tyi, main = "Distribution of type of injury")
tab1(kbbh$mot, main = "Distribution of Mode of transport")
tab1(kbbh$RTS, main = "Distribution of RTS")
tab1(kbbh$hd, main = "Distribution of hospital disposition")
```
