---
title: Profile of patients triaged green in the Emergency Department (ED) of a secondary care hospital in Mumbai
author: "Anna"
date: "12/05/2020"
output:
  word_document: default
  html_document: 
    fig_caption: yes
    fig_height: 3
    fig_width: 5
---


```{r setup, include=FALSE, results='asis', message=FALSE}
library(knitr)
library(dplyr)
library(psych)
library(data.table) 
library(tableone)
library(epiDisplay)
library(tigerstats)
library(splitstackshape)
library(tidyverse)
library(tidyr)
library(assertthat)
library(rlist)
opts_chunk$set(echo = FALSE)
```

# Abstract

## Introduction

Trauma is one of the leading causes of emergency department visits.  Based on injury severity and physiological parameters patients are triaged as red, orange, yellow, or green, from most to least urgent. Emergency departments provide quick and easy access to healthcare, leading to visits of patients with no injury or minor injuries. Although these patients are triaged green they consume significant resources. It is important to understand the characteristics of these patients to understand how they should be managed so that resources can be directed towards serious patients.

## Aim

The aim of this study is to describe the demographic, physiological and injury characteristics of trauma patients triaged green in the ED of a secondary healthcare hospital in Mumbai.

## Methods

Data of patients triaged green were extracted from the Trauma Triage Study in India dataset. A retrospective cohort analysis of the data was performed using R. 

## Results
```{r Data_cleaning_manipulation, warning=FALSE, results='hide'}

# Get data directory

data.dir <- getOption("data.dir")
if (is.null(data.dir))
    stop ("Please set the directory where the data is kept using options(data.dir = \"<path to data directory>\")")

## datatable with ISS
g <- data.table::fread(file.path(data.dir, "ttris-dataset-with-iss-8317-20190708120045.csv"))
g1 <- g %>% dplyr::select(pid,X0:iss)

## datatable without ISS
h <- data.table::fread(file.path(data.dir, "TTRIS.csv"))

## Merged the two datatables
h$pid <- as.integer(h$pid)
dt <- merge(g1, h, by = "pid") # When you do this you loose all extra patients present in g mot not in h. Is that intentional? Yes, g has the AIS score columns.

## Rearranging columns
dt <- dt %>% dplyr::select(centre,pid,cr,qcrec:vsqs,ic:toa,egcs:rr,nsi,tc,dvsr:tvsr,dus:testvar,locked,X0:iss,doi_toi:valid)

extra_col <- dt %>% dplyr::select(doi_toi:valid)

## Adding the extra columns to g
g <- cbind(g,extra_col) # So here you are putting in data in g from dt. Why do you want to do that? We need same number of columns in both the dataset to be able to change the class, along with them being in the same order.

## Changing class of the data
g <- as.data.frame(g)
dt <- as.data.frame(dt)

dt[] <- lapply(names(dt), function(name) {
    column <- dt[, name]
    correct.class <- class(g[, name])
    if (correct.class == "factor") {
        new.column <- factor(column, levels = levels(g[, name]))
    } else {
        new.column <- as(column, correct.class)
    }
    return (new.column)
})

## dplyr::all_equal(g,dt) ##To compare the two dataframes

# Creating age groups
agebreaks <- c(17,25,45,65,98)
agelabels <- c("18-24","25-44","45-64","65+")
dt$agegrp <-cut(dt$age,
                breaks = agebreaks,
                right = FALSE,
                labels = agelabels)

# Mechanism of Injury
moi.collapsed <- moi <- as.character(dt$moi)

get_icd_range <- function(first.char, start.number, end.number) {
    icd.range <- paste0(
        paste0(
            "^",
            first.char,
            stringr::str_pad(start.number:end.number, 2, pad = "0"),
            "[0-9]?$"),
        collapse = "|")
    return (icd.range)
}

icd.ranges <- list(c("Transport accident" = get_icd_range("V", 0, 99)),
                   c("Fall" = get_icd_range("W", 0, 19)),
                   c("Animal bite" = get_icd_range("W", 50, 64)),
                   c("Assault" = paste0(get_icd_range("X", 85, 99), "|", get_icd_range("Y", 0, 9))))

for(icd.range in icd.ranges) moi.collapsed[grep(icd.range, moi)] <- names(icd.range)
moi.collapsed[!(moi.collapsed %in% sapply(icd.ranges, attr, "names"))] <- "Other"
dt$moi.collapsed <- as.factor(moi.collapsed)

# Correcting ISS
dt <- as_tibble(dt)
ais <- dt %>% dplyr:: select(starts_with("clean_"))

ais.list <- lapply(names(ais), function(x) cSplit(ais[, x], x, sep = ","))
row_max <- function(x, na.rm = FALSE) {
    if (all(is.na(x)))
        return (NA)
    max(x, na.rm = na.rm)
}
ais.max <- lapply(ais.list, function(x) apply(x, 1, row_max, na.rm = TRUE))
get_top_3 <- function(x) {
    x <- sort(x, decreasing = TRUE, na.last = TRUE)
    i <- min(c(3, length(x)))
    x[seq_len(i)]
}
calculate_iss <- function(x) {
    top3 <- get_top_3(x)
    iss <- sum(top3[1]^2, top3[2]^2, top3[3]^2, na.rm = TRUE)
    if (all(is.na(top3)))
        iss <- NA
    iss
}
iss <- apply(do.call(cbind, ais.max), 1, calculate_iss)
iss[is.na(iss)] <- "0"

dt$iss <- as.numeric(iss)

## Note that these operations do not work as intended because iss is
## character, not numeric

issbreaks <- c(0,1,9,16,25,109)
isslabels <- c("ND","Minor","Moderate","Severe","Profound")
dt$isspv <-cut(dt$iss,
                breaks = issbreaks,
                right = FALSE,
                labels = isslabels)

# Append a new variable 'GCS' to the dataset
gcs <- with(dt, egcs + vgcs + mgcs)
gcs <- replace(gcs, gcs > 15, NA)

## What's happening here? Assigning parameter value for calculating RTS. Something wrong?

gcspv <- gcs
gcspv[gcs >= 13] <- 4
gcspv[gcs >= 9 & gcs < 13] <- 3
gcspv[gcs >= 6 & gcs < 9] <- 2
gcspv[gcs >= 4 & gcs < 6] <- 1
gcspv[gcs < 4] <- 0

## RR parameter value for calculating RTS
rr <- with(dt,replace(rr, rr == 999, NA))
rrpv <- rr

rrpv[rr >= 10 & rr <= 29] = 4
rrpv[rr > 29] = 3
rrpv[rr >= 6 & rr <= 9] = 2
rrpv[rr >= 1 & rr <= 5] = 1
rrpv[rr == 0] = 0

## SBP parameter value for calculating RTS
sbp <- with(dt,replace(sbp, sbp == 999, NA))
sbppv <- sbp

sbppv[sbp > 89] <- 4
sbppv[sbp >=76 & sbp <= 89] <- 3
sbppv[sbp >= 50 & sbp <= 75] <- 2
sbppv[sbp >= 1 & sbp <= 49] <- 1
sbppv[sbp == 0] <- 0

## Formula for calculating RTS

dt$rts = round(0.9368*as.numeric(gcspv) + 
               0.7326*as.numeric(sbppv) + 
               0.2908*as.numeric(rrpv), digits = 0)

# Complete case analysis
dt$s24h <- replace(dt$s24h, dt$s24h == 999, 888) ##Loss of follow up 
dt$s30d <- replace(dt$s30d, dt$s30d == 999, 888)
dt$s6m <- replace(dt$s6m, dt$s6m == 999, 888)
dt$s <- replace(dt$s, dt$s == 999, 888)

## Converting all Injury 999 to Not Defined
inj_cols <- dt[,grep("inj", colnames(dt))]

inj_cols <- inj_cols %>% 
    mutate_all(funs(str_replace(., "999", "ND")))

dt[,grep("inj", colnames(dt))] <- inj_cols

## Converting NA to ND in max cols
#dt %>% dplyr:: select(starts_with("max_")) %>% 
#    mutate_all(funs(str_replace(., "NA", "ND")))

## Dropping unwanted columns
drop.cols2 <- dt %>% dplyr:: select(starts_with("clean_"), starts_with("max_"), qcrec, vsqs, testvar, locked, X0)

drop.cols2 <- colnames(drop.cols2)

dt <- dt[, !(colnames(dt) %in% c(drop.cols2))]

## Converting rest of the 999s to NA
dt[] <- dt[] %>% 
    mutate_all(funs(na_if(.,"999")))

missing.data.list <- lapply(dt, function(column) {
    n.na <- sum(is.na(column))
    p.na <- round((n.na/length(column)) * 100)
    missing <- data.frame("Count" = n.na, "Percentage" = p.na)
    return (missing)
})

missing.data <- do.call(rbind, missing.data.list)

maximum.missing <- rownames(missing.data)[missing.data$Count == max(missing.data$Count)]

complete.index <- complete.cases(dt)
n.incomplete <- sum(!complete.index)
p.incomplete <- round((n.incomplete/nrow(dt)) * 100)

dtt <- dt[complete.index, ]

```

```{r result}
# Created a dataframe of triage category green
kbbh <- subset(x=dtt,subset = dtt$tc == 0)

# Number of records
dm <- as.numeric(nrow(kbbh))

# Age
mean_age <- round(mean(kbbh$age),2)
stddev_age <- round(sd(kbbh$age),2)

nm_age <- sum(kbbh$agegrp != "65+")
percent_agebelow65 <- round(mean(kbbh$agegrp != "65+") * 100, 2)

calculate_n_p <- function(level.name, factor.var){
    assert_that(is.character(level.name))
    assert_that(is.factor(factor.var))
    nm <- sum(factor.var == level.name)
    percent <- round((nm/dm) * 100,2)
    return.list <-list(nm = nm, percent = percent)
    return.list
}

# Sex
sex.data <- list.sort(lapply(setNames(nm = levels(kbbh$sex)), calculate_n_p, factor.var = kbbh$sex), nm, decreasing = TRUE)

sexv <- unlist(sex.data)
nm_male <- sexv["Male.nm"]
p_male <- sexv["Male.percent"]
nm_female <- sexv["Female.nm"]
p_female <- sexv["Female.percent"]

# Type of Injury
nm_blunt <- sum(kbbh$tyi == 0)
percent_blunt <- round(mean(kbbh$tyi == 0) * 100,2)

# Mode of Transport
mot.data <- list.sort(lapply(setNames(nm = levels(kbbh$mot)), calculate_n_p, factor.var = kbbh$mot), nm, decreasing = TRUE)

motv <- unlist(mot.data)
nm_pvt <- motv["Private Vehicle.nm"]
p_mot_pvt <- motv["Private Vehicle.percent"]
nm_mot_onfoot <- motv["On foot.nm"]
p_mot_onfoot <- motv["On foot.percent"]
mot1 <- names(mot.data[1])

# MOI
moi.data <- list.sort(lapply(setNames(nm = levels(kbbh$moi.collapsed)), calculate_n_p, factor.var = kbbh$moi.collapsed), nm, decreasing = TRUE)

moiv <- unlist(moi.data)
p_animalbite <- moiv["Animal bite.percent"]
p_assualt <- moiv["Assualt.percent"]
p_fall <- moiv["Fall.percent"]
p_other <- moiv["Other.percent"]
p_tra <- moiv["Transport accident.percent"]

moi1 <- names(moi.data[1])
moi2 <- names(moi.data[2])
moi3 <- names(moi.data[3])

# Outcome
kbbh$s6m <- as.factor(kbbh$s6m)
outcome.data <- lapply(setNames(nm = levels(kbbh$s6m)), calculate_n_p, factor.var = kbbh$s6m)
p_outcome_alive <- outcome.data[["0"]]["percent"]
p_outcome_dead <- outcome.data[["1"]]["percent"]
n_outcome_dead <- outcome.data[["1"]]["nm"]

p_followup <- as.numeric(p_outcome_alive)+as.numeric(p_outcome_dead)

# Grouping patients wrt to injury

crush_inj <- c("S07","S17","S280","S380","S381","T041","S47","S57","S67","T042","S770","S771","S772","S87","S97","T043","T048","T049")

internal_inj <- c("S06","T905","S140","S141","S240","S241","S340","S341","S343","T093","T913","S142","S260","S270","S271","S272","S273","S274","S275","S276","S278","S279","T914","S36","S37","S396","S397","T065","T915")

bv_inj <- c("S090","S150","S512","S153","S154","S155","S156","S157","S158","S159","S151","S25","S350","S351","S352","S353","S354","S355","S357","S358","S359","S45","S55","S65","T114","S75","S85","S95","T134","T063","T145")

amp_inj <- c("S081","S082","S083","S084","S085","S086","S087","S088","S089","S18","S281","S382","S383","T096","S48","S58","S68","T050","T052","S780","T058","T059")

frac_inj <- c("S020","S021","S023","S027","S029","T902","S022","S024","S025","S026","S128","S129","S120","S121","S122","S123","S124","S125","S126","S127","S220","S221","S320","S321","S322","T08","T911","S222","S223","S224","S225","S226","S227","S228","S229","S323","S324","S325","S326","S327","S328","T021","T912","S42","S52","S62","T022","T024","T10","T921","T922","S720","S721","S722","S723","S")

apply(kbbh, 1, function(x) if(any(x %in% crush_inj)) cat("Crush Injury"))
apply(kbbh, 1, function(x) if(any(x %in% internal_inj)) cat("Internal organ Injury"))
apply(kbbh, 1, function(x) if(any(x %in% bv_inj)) cat("Blood Vessel Injury"))
apply(kbbh, 1, function(x) if(any(x %in% amp_inj)) cat("Amputation"))

```

There was `r nrow(dt)` observations in the study data. Out of these, `r n.incomplete` (`r p.incomplete` %) observations were incomplete.

In the TTRIS dataset, `r dm` patients were triaged green. The mean age of these patients was `r mean_age` (+-`r stddev_age`). `r percent_agebelow65`% of the patients were below 65 years of age. `r p_male`% patients were male, while `r p_female`% patients were female. `r percent_blunt`% patients had blunt injuries. Most common mode of arrival to the ED was through `r mot1` (`r p_mot_pvt`)%. `r p_mot_onfoot`% patients came to the ED on foot. `r moi1` were the most common cause of injury (`r p_assualt`%) followed by `r moi2` (`r p_tra`%) and `r moi3` (`r p_fall`%). Injury due to animal bites were `r p_animalbite`. `r p_followup`% patients were successfully followed up at 6months, of which `r n_outcome_dead` patients were reported dead.

```{r tables}

Table1 <- colPerc(xtabs(~kbbh$isspv+kbbh$moi.collapsed, data = kbbh))
kable(Table1, 
      caption = "Table 1. Percentage distribution of Injury severity among different Mechanism of Injury",
)

```

## Conclusion

Among the patients triaged green, injury due to assualt and transport accidents were predominant. `r p_rts_8`% patients had an RTS score of 8. Patients with minor injuries and no injuries defined are `r p_iss_mi`% and `r p_iss_nd`% respectively. 

# Introduction
Trauma is one of the leading causes of emergency department visits. 

Based on injury severity and physiological parameters patients are triaged as red, orange, yellow, or green, from most to least urgent. 

Emergency departments provide quick and easy access to healthcare, leading to visits of patients with no injury or minor injuries. 

Although these patients are triaged green they consume significant resources.

It is important to understand the characteristics of these patients to understand how they should be managed so that resources can be directed towards serious patients.

The aim of this study is to describe the demographic, physiological and injury characteristics of trauma patients triaged green in the ED of a secondary healthcare hospital in Mumbai.

# Methods

Data of patients triaged green were extracted from the Trauma Triage Study in India dataset. A retrospective cohort analysis of the data was performed using R.

# Results

There was `r nrow(dt)` observations in the study data. Out of these, `r n.incomplete` (`r p.incomplete` %) observations were incomplete.

In the TTRIS dataset, `r dm` patients were triaged green. The mean age of these patients was `r mean_age` (+-`r stddev_age`). `r percent_agebelow65`% of the patients were below 65 years of age. `r percent_male`% patients were male, while `r percent_female`% patients were female. `r percent_blunt`% patients had blunt injuries. Most common mode of arrival to the ED was through `r mot1` (`r percent_mot_pvt`)%. `r percent_mot_onfoot`% patients came to the ED on foot. `r moi1` were the most common cause of injury (`r p_assualt`%) followed by `r moi2` (`r p_tra`%) and `r moi3` (`r p_fall`%). Injury due to animal bites were `r p_animalbite`. `r p_followup`% patients were successfully followed up at 6months, of which `r n_outcome_dead` patients were reported dead.

```{r}
#Factor labels
kbbh$sex <- factor(kbbh$sex, levels = c(0,1), labels = c("Female","Male"))
kbbh$tyi <- factor(kbbh$tyi, levels = c(0,1,2), 
                   labels = c("Blunt","Penetrating","Blunt & penetrating"))
kbbh$mot <- factor(kbbh$mot, levels = c(0,1,2,3), 
                   labels = c("Ambulance","Police","Private Vehicle","On foot"))
kbbh$tran <- factor(kbbh$tran, levels = c(0,1), 
                    labels = c("Direct","Transferred"))
kbbh$avpu <- factor(kbbh$avpu, levels = c(0,1,2,3), 
                    labels = c("Unresponsive","Pain","Verbal","Alert"))
kbbh$vo2 <- factor(kbbh$vo2, levels = c(0,1), 
                   labels = c("Not on oxygen support","On oxygen support"))
kbbh$nsi <- factor(kbbh$nsi, levels = c(0,1,2), 
                   labels = c("No serious injury","Single","Multiple"))
kbbh$hd <- factor(kbbh$hd, levels = c(0,1,2,3), 
                  labels = c("Discharged","Dead","Transferred","Abscond/ DAMA"))

myvars <- c("age", "agegrp", "sex", "moi.collapsed", "tyi", "mot",
"tran", "avpu", "sbp", "dbp", "hr", "spo2", "vo2", "rr", "nsi", "hd", "gcs", "isspv", "rts")

catvars <- c("sex", "moi.collapsed", "tyi", "mot", "tran", "avpu", "vo2", "nsi", "hd", "gcs", "rts")

Table2 <- CreateTableOne(vars = myvars, data=kbbh, factorVars = catvars)
knitr::kable(print(Table2, caption = "Table 2. Demographic, Physiological parameters, Injury Characteristics", showAllLevels = TRUE, printToggle = FALSE))

p_iss_nd <- round(mean(kbbh$isspv == "ND") * 100,2)
p_iss_mi <- round(mean(kbbh$isspv == "Minor") * 100,2)
p_rts_8 <- round(mean(kbbh$rts == 8) * 100,2)

```

