---
title: Profile of patients triaged green in the Emergency Department (ED) of a secondary care hospital in Mumbai
author: "Anna"
date: "12/05/2020"
output:
  word_document: default
  html_document: 
    fig_caption: yes
    fig_height: 3
    fig_width: 5
---


```{r setup, include=FALSE, results='asis', message=FALSE}
library(knitr)
library(dplyr)
library(psych)
library(data.table) 
library(tableone)
library(epiDisplay)
library(tigerstats)
library(splitstackshape)
library(tidyverse)
library(tidyr)
opts_chunk$set(echo = FALSE)
```

# Abstract

## Introduction

Injuries are one of the leading causes of Emergency Department visits. Physicians triage patients into red, orange, yellow and green, from most to least urgent based on injruy severity and physiological parameters. The proportion of patients that are triaged green is larger compared with patients with life-threatning injuries and injuries that require urgent attention<!-- This sentence is a bit awkward. Do you mean that more patients are triage green compared with red?-->. It is important to understand the characteristics of the patients triaged green to understand how they should be managed to allow resources to be directed towards serious patients.

## Aim

The aim of this study is to describe the demographic, physiological and injury characteristics of trauma patients triaged green in the ED of a secondary healthcare hospital in Mumbai.

## Methods

Data of patients triaged green were extracted from the Trauma Triage Study in India dataset. A retrospective cohort analysis of the data was performed using R. 

## Results
```{r Data_cleaning, warning=FALSE, results='hide'}

# read the original csv file

## Get data directory

options(data.dir = "C:\\Users\\annan\\OneDrive\\Documents\\Projects\\PGTP")

data.dir <- getOption("data.dir")
if (is.null(data.dir))
    stop ("Please set the directory where the data is kept using options(data.dir = \"<path to data directory>\")")

## datatable with ISS
g <- data.table::fread(file.path(data.dir, "ttris-dataset-with-iss-8317-20190708120045.csv"))

## datatable without ISS
h <- data.table::fread(file.path(data.dir, "TTRIS.csv"))

## Merged the two datatables

h$pid <- as.integer(h$pid)

merged <- merge(g, h, by = "pid")

## Duplicated rows
merged <- merged[!duplicated(merged$pid), ]

## Changed column names
names(merged)[names(merged) == "iss.y"] <- "delcolumn.x"
names(merged)[names(merged) == "iss.x"] <- "iss.y"

## Removing duplicated columns
dt <- data.table(merged)
drop.cols <- grep(".x$", colnames(dt))
dt[, (drop.cols) := NULL]

## Removed .y from the coulmn names
names(dt) <- sub('\\.[^.]+$', "", names(dt))

## Rearranging columns

#dt <- dt %>% dplyr::select(order(colnames(.)))

dt <- dt %>% dplyr::select(centre,pid,cr,qcrec:vsqs,ic:toa,egcs:rr,nsi,tc,dvsr:tvsr,dus:testvar,locked,X0:iss,doi_toi:valid)

extra_col <- dt %>% dplyr::select(doi_toi:valid)

## Adding the extra columns to g

g <- cbind(g,extra_col)

## Changing class of the columns

g <- as.data.frame(g)
dt <- as.data.frame(dt)

dt[] <- lapply(names(dt), function(name) {
    column <- dt[, name]
    correct.class <- class(g[, name])
    if (correct.class == "factor") {
        new.column <- factor(column, levels = levels(g[, name]))
    } else {
        new.column <- as(column, correct.class)
    }
    return (new.column)
})

## dplyr::all_equal(g,dt) ##To compare the two dataframes

# Calculating ISS

dt <- as.tibble(dt)

ais <- dt %>% dplyr:: select(starts_with("clean_"))

ais.list <- lapply(names(ais), function(x) cSplit(ais[, x], x, sep = ","))
row_max <- function(x, na.rm = FALSE) {
    if (all(is.na(x)))
        return (NA)
    max(x, na.rm = na.rm)
}
ais.max <- lapply(ais.list, function(x) apply(x, 1, row_max, na.rm = TRUE))
get_top_3 <- function(x) {
    x <- sort(x, decreasing = TRUE, na.last = TRUE)
    i <- min(c(3, length(x)))
    x[seq_len(i)]
}
calculate_iss <- function(x) {
    top_3 <- get_top_3(x)
    iss <- sum(x[1]^2, x[2]^2, x[3]^2, na.rm = TRUE)
    iss
}
iss <- apply(do.call(cbind, ais.max), 1, calculate_iss)
dt$iss <- iss

# Complete case analysis

dt$s24h <- replace(dt$s24h, dt$s24h == 999, 99) ##Loss of follow up 
dt$s30d <- replace(dt$s30d, dt$s30d == 999, 99)
dt$s6m <- replace(dt$s6m, dt$s6m == 999, 99)
dt$s <- replace(dt$s, dt$s == 999, 99)

## Converting all Injury 999 to Not Defined

inj_cols <- dt[,grep("inj", colnames(dt))]

inj_cols <- inj_cols %>% 
    mutate_all(funs(str_replace(., "999", "ND")))

dt[,grep("inj", colnames(dt))] <- inj_cols

## Converting NA to ND in max cols

dt %>% dplyr:: select(starts_with("max_")) %>% 
    mutate_all(funs(str_replace(., "NA", "ND")))

## Dropping unwanted columns
drop.cols2 <- dt %>% dplyr:: select(starts_with("clean_"), starts_with("max_"), qcrec, vsqs, testvar, locked, X0)

drop.cols2 <- colnames(drop.cols2)

dt <- dt[, !(colnames(dt) %in% c(drop.cols2))]

## Converting rest of the 999s to NA
dt[] <- dt[] %>% 
    mutate_all(funs(na_if(.,"999")))

missing.data.list <- lapply(dt, function(column) {
    n.na <- sum(is.na(column))
    p.na <- round((n.na/length(column)) * 100)
    missing <- data.frame("Count" = n.na, "Percentage" = p.na)
    return (missing)
})

missing.data <- do.call(rbind, missing.data.list)

maximum.missing <- rownames(missing.data)[missing.data$Count == max(missing.data$Count)]

complete.index <- complete.cases(dt)
n.incomplete <- sum(!complete.index)
p.incomplete <- round((n.incomplete/nrow(dt)) * 100)

dtt <- dt[complete.index, ]

```


```{r age_sex_toi_mot}

# Created a dataframe of triage category green
kbbh <- subset(x=dtt,subset = dtt$tc == 0)

# Number of records
dm <- as.numeric(nrow(kbbh))

# Age
kbbh$age <- as.numeric(kbbh$age)
kbbh$age <- replace(kbbh$age, kbbh$age == 999, NA)
mean_age <- round(mean(kbbh$age, na.rm = TRUE),2)
stddev_age <- round(sd(kbbh$age, na.rm = TRUE),2)

## Creating age groups
agebreaks <- c(17,25,45,65,98)
agelabels <- c("18-24","25-44","45-64","65+")
setDT(kbbh)[,agegrp := cut(kbbh$age,
                           breaks = agebreaks,
                           right = FALSE,
                           labels = agelabels)]

## Calculating percentage of patients below age 65

dm_age <- nrow(subset(kbbh, !is.na(age))) 

age_below65 <- dplyr::filter(kbbh, kbbh$agegrp != "65+") 
nm_age <- nrow(age_below65) # neumerator
percent_agebelow65 <- round((nm_age/ dm) * 100, 2)

## The above could have been achieved with
nm_age <- sum(kbbh$agegrp != "65+", na.rm = TRUE)
percent_agebelow65 <- round(mean(kbbh$agegrp != "65+", na.rm = TRUE) * 100, 2)

# Sex

kbbh$sex <- as.numeric(kbbh$sex)

male <- dplyr::filter(kbbh, kbbh$sex == 1) # neumerator
nm_male <- nrow(male)
percent_male <- round((nm_male/ dm) * 100,2)

female <- dplyr::filter(kbbh, kbbh$sex == 0) # neumerator
nm_female <- nrow(female)
percent_female <- round((nm_female/ dm) * 100,2)

# Type of Injury

kbbh$tyi <- as.numeric(kbbh$tyi)

blunt <- dplyr::filter(kbbh, kbbh$tyi == 0) # neumerator
nm_blunt <- nrow(blunt)
percent_blunt <- round((nm_blunt/ dm) * 100,2)

# Mode of Transport

kbbh$mot <- as.numeric(kbbh$mot)

kbbh$mot <- replace(kbbh$mot, kbbh$mot == 999, NA)

kbbh$mot <- replace(kbbh$mot, kbbh$mot == 2, "Private Vehicles")
mot_pvt <- dplyr::filter(kbbh, kbbh$mot == "Private Vehicles") # neumerator
nm_pvt <- nrow(mot_pvt)
percent_mot_pvt <- round((nm_pvt/ dm) * 100,2)

sort_most_common <- function(v) {
   uniqv <- unique(v)
   uniqv[sort.list(tabulate(match(v, uniqv)), decreasing = TRUE)]
}

sorted_mot <- sort_most_common(kbbh$mot)

mot1 <- sorted_mot[[1]][]

mot_onfoot <- dplyr::filter(kbbh, kbbh$mot == 3) # neumerator
nm_onfoot <- nrow(mot_onfoot)
percent_mot_onfoot <- round((nm_onfoot/ dm) * 100,2)

```


```{r moi_followups, results='hide'}

# Mechanism of Injury
moi.collapsed <- moi <- as.character(kbbh$moi)

get_icd_range <- function(first.char, start.number, end.number) {
    icd.range <- paste0(
        paste0(
            "^",
            first.char,
            stringr::str_pad(start.number:end.number, 2, pad = "0"),
            "[0-9]?$"),
        collapse = "|")
    return (icd.range)
}

icd.ranges <- list(c("Transport accident" = get_icd_range("V", 0, 99)),
                   c("Fall" = get_icd_range("W", 0, 19)),
                   c("Animal bite" = get_icd_range("W", 50, 64)),
                   c("Assault" = paste0(get_icd_range("X", 85, 99), "|", get_icd_range("Y", 0, 9))))

for(icd.range in icd.ranges) moi.collapsed[grep(icd.range, moi)] <- names(icd.range)
moi.collapsed[!(moi.collapsed %in% sapply(icd.ranges, attr, "names"))] <- "Other"

kbbh$moi.collapsed <- as.factor(moi.collapsed)

calculate_n_p_moi <- function(level.name) {
    moi.level <- dplyr::filter(kbbh, kbbh$moi.collapsed == level.name) # neumerator
    nm <- nrow(moi.level)
    percent <- round((nm/dm) * 100,2)
    return.list <- list(nm = nm, percent = percent)
    return (return.list)
}

moi.data <- lapply(setNames(nm = levels(kbbh$moi.collapsed)), calculate_n_p_moi)

p_animalbite <- moi.data[[1]][2]
p_assualt <- moi.data[[2]][2]
p_fall <- moi.data[[3]][2]
p_other <- moi.data[[4]][2]
p_tra <- moi.data[[5]][2]

sorted_moi <- sort_most_common(kbbh$moi.collapsed)

moi1 <- sorted_moi[[1]][]
moi2 <- sorted_moi[[2]][]
moi3 <- sorted_moi[[3]][]

# Trauma outcomes

kbbh$s6m <- as.numeric(kbbh$s6m)

kbbh$s6m <- as.factor(kbbh$s6m)

calculate_n_p_outcome <- function(level.name) {
    outcome.level <- dplyr::filter(kbbh, kbbh$s6m == level.name) # neumerator
    nm <- nrow(outcome.level)
    percent <- round((nm/dm) * 100,2)
    return.list <- list(nm = nm, percent = percent)
    return (return.list)
}

outcome.data <- lapply(setNames(nm = levels(kbbh$s6m)), calculate_n_p_outcome)

str(outcome.data)

p_outcome_alive <- outcome.data[[1]][2]
p_outcome_dead <- outcome.data[[2]][2]
n_outcome_dead <- outcome.data[[2]][1]

## Successful Followups
percent_followup <- as.numeric(p_outcome_alive)+as.numeric(p_outcome_dead)

```

There was `r nrow(dt)` observations in the study data. Out of these, `r n.incomplete` (`r p.incomplete` %) observations were incomplete.

In the TTRIS dataset, `r dm` patients were triaged green. The mean age of these patients was `r mean_age` (+-`r stddev_age`). `r percent_agebelow65`% of the patients were below 65 years of age. `r percent_male`% patients were male, while `r percent_female`% patients were female. `r percent_blunt`% patients had blunt injuries. Most common mode of arrival to the ED was through `r mot1` (`r percent_mot_pvt`)%. `r percent_mot_onfoot`% patients came to the ED on foot. `r moi1` were the most common cause of injury `r p_assualt`% followed by `r moi2` `r p_tra`% and `r moi3` `r p_fall`%. Injury due to animal bites were. `r percent_followup` patients were successfully followed up at 6months, of which `r n_outcome_dead` patients were reported dead. 

```{r ISS_category}

kbbh$iss <- as.numeric(kbbh$iss)

# Categorizing ISS Score

kbbh$isspv[kbbh$iss >= 16 & kbbh$iss <= 24] = 'Severe'
kbbh$isspv[kbbh$iss >= 9 & kbbh$iss <= 15] = 'Moderate'
kbbh$isspv[kbbh$iss == 1 & kbbh$iss <= 8] = 'Minor'
kbbh$isspv[is.na(kbbh$iss)] = 'No injuries defined'

Table2 <- colPerc(xtabs(~kbbh$isspv+kbbh$moi.collapsed, data = kbbh))
kable(Table2, 
      caption = "Table 1. Percentage distribution of Injury severity among different Mechanism of Injury",
)

```

```{r RTS}

## Append a new variable 'GCS' to the dataset

kbbh$egcs <- as.numeric(kbbh$egcs)
kbbh$vgcs <- as.numeric(kbbh$vgcs)
kbbh$mgcs <- as.numeric(kbbh$mgcs)

gcs <- with(kbbh, egcs + vgcs + mgcs)

gcs <- replace(gcs, gcs > 15, NA)

gcspv <- gcs
gcspv[gcs >= 13] <- 4
gcspv[gcs >= 9 & gcs < 13] <- 3
gcspv[gcs >= 6 & gcs < 9] <- 2
gcspv[gcs >= 4 & gcs < 6] <- 1
gcspv[gcs < 4] <- 0

## RR parameter value for calculating RTS

kbbh$rr <- as.numeric(kbbh$rr)

rr <- with(kbbh,replace(rr, rr == 999, NA))

rrpv <- rr

rrpv[rr >= 10 & rr <= 29] = 4
rrpv[rr > 29] = 3
rrpv[rr >= 6 & rr <= 9] = 2
rrpv[rr >= 1 & rr <= 5] = 1
rrpv[rr == 0] = 0

## SBP parameter value for calculating RTS

kbbh$sbp <- as.numeric(kbbh$sbp)

sbp <- with(kbbh,replace(sbp, sbp == 999, NA))

sbppv <- sbp
sbppv[sbp > 89] <- 4
sbppv[sbp >=76 & sbp <= 89] <- 3
sbppv[sbp >= 50 & sbp <= 75] <- 2
sbppv[sbp >= 1 & sbp <= 49] <- 1
sbppv[sbp == 0] <- 0

## Formula for calculating RTS

## Why do you want to round it using 0 digits?
## So that I get a whole number.The decimals make a longer table.

kbbh$RTS = round(0.9368*as.numeric(gcspv) + 
                 0.7326*as.numeric(sbppv) + 
                 0.2908*as.numeric(rrpv), digits = 0)

Table2 <- rowPerc(xtabs(~kbbh$RTS, data = kbbh))
kable(Table2, 
      caption = "Table 1. RTS of patients",
)

```

## Conclusion

Among the patients triaged green, injury due to assualt and transport accidents were predominant.98.94% patients had an RTS score of 8. Patients with minor injuries and no injuries defined are upto 99.75%. 

## Analysis

```{r Extra, warning=FALSE, results='hide'}

## Table of age group
tab1(kbbh$agegrp, main = "Distribution of patients by Age")

tab1(kbbh$sex, main = "Distribution of patients by Gender")

kbbh$tyi <- replace(kbbh$tyi, kbbh$tyi == 999, NA)

tab1(kbbh$tyi, main = "Distribution of type of injury")

kbbh$mot <- replace(kbbh$mot, kbbh$mot == 999, NA)

tab1(kbbh$mot, main = "Distribution of Mode of transport")

tab1(kbbh$RTS, main = "Distribution of RTS")

tab1(kbbh$hd, main = "Distribution of hospital disposition")

```
