---
title: Profile of patients triaged green in the Emergency Department (ED) of a secondary care hospital in Mumbai
author: "Anna"
date: "12/05/2020"
output:
  word_document: default
  html_document: 
    fig_caption: yes
    fig_height: 3
    fig_width: 5
---


```{r setup, include=FALSE, results='asis', message=FALSE}
library(knitr)
library(dplyr)
library(psych)
library(data.table)
library(tableone)
library(epiDisplay)
library(tigerstats)
opts_chunk$set(echo = FALSE)
```

# Abstract

## Introduction

Injuries are one of the leading causes of Emergency Department visits. Physicians triage patients into red, orange, yellow and green, from most to least urgent based on injruy severity and physiological parameters. The proportion of patients that are triaged green is larger compared with patients with life-threatning injuries and injuries that require urgent attention<!-- This sentence is a bit awkward. Do you mean that more patients are triage green compared with red?-->. It is important to understand the characteristics of these patients <!-- It's not obvious that "these patients" here refers to the patients triaged green. Try to make that clearer.-->to know if there is a need for them to be managed differently so that the manpower and resources of the emergency department can be directed towards serious patients<!-- This sentence is a bit long. Try to be more direct: "... to understand how they should be managed to allow resources to be directed towards serios patients", or something like that.-->

## Aim

The aim of this study is to describe the demographic, physiological and injury characteristics of trauma patients triaged green in the ED of a secondary healthcare hospital in Mumbai.

## Methods

Data of patients triaged green were extracted from the Trauma Triage Study in India dataset. A retrospective cohort analysis of the data was performed using R. 

## Results
```{r, warning=FALSE, results='hide'}

# read the original csv file

## Now let's take this one level further. You ideally don't want to
## specify any file paths here, because that means that the next
## person, like me, have to change the file path directly here to get
## it to work. Let's add a call to options instead.

## Get data directory
data.dir <- getOption("data.dir")
if (is.null(data.dir))
    stop ("Please set the directory where the data is kept using options(data.dir = \"<path to data directory>\")")

## datatable with ISS
g <- data.table::fread(file.path(data.dir, "ttris-dataset-with-iss-8317-20190708120045.csv"))

## datatable without ISS
h <- data.table::fread(file.path(data.dir, "TTRIS.csv"))

#h$iss <- NA

#h[] <- mapply(FUN = as,h,sapply(g,class),SIMPLIFY = FALSE)


## Merged the two datatables

h$pid <- as.integer(h$pid)

merged <- merge(g, h, by = "pid")

## Note that you have a number of duplicate pids in your merged data
merged$pid[duplicated(merged$pid)]

## Changed column names
names(merged)[names(merged) == "iss.y"] <- "delcolumn.x"
names(merged)[names(merged) == "iss.x"] <- "iss.y"

## Removing duplicated columns
dt <- data.table(merged)
drop.cols <- grep(".x$", colnames(dt))
dt[, (drop.cols) := NULL]

## Removed .y from the coulmn names

names(dt) <- sub(".y$", "", names(dt))
make.names(names(dt)) # What are you trying to achieve here? This will only print the output to the console, but maybe that's what you're after?

```


```{r}

# Created a dataframe of triage category green
kbbh <- subset(x=dt,subset = dt$tc == 0)

# Number of records
dm <- as.numeric(nrow(kbbh))

# Age
kbbh$age <- as.numeric(kbbh$age)
kbbh$age <- replace(kbbh$age, kbbh$age == 999, NA)
mean_age <- round(mean(kbbh$age, na.rm = TRUE),2)
stddev_age <- round(sd(kbbh$age, na.rm = TRUE),2)

## Creating age groups
agebreaks <- c(17,25,45,65,98)
agelabels <- c("18-24","25-44","45-64","65+")
setDT(kbbh)[,agegrp := cut(kbbh$age,
                           breaks = agebreaks,
                           right = FALSE,
                           labels = agelabels)]

## Calculating percentage of patients below age 65

#dm_age <- nrow(subset(kbbh, !is.na(age))) 

age_below65 <- dplyr::filter(kbbh, kbbh$agegrp != "65+") 
nm_age <- nrow(age_below65) # neumerator
percent_agebelow65 <- round((nm_age/ dm) * 100, 2) ## This includes those with missing age in the denominator. Is that intentional? 

## The above could have been achieved with
nm_age <- sum(kbbh$agegrp != "65+", na.rm = TRUE)
percent_agebelow65 <- round(mean(kbbh$agegrp != "65+", na.rm = TRUE) * 100, 2)

# Sex
#dm_sex <- nrow(subset(kbbh, !is.na(sex))) # denominator if excluding NA

kbbh$sex <- as.numeric(kbbh$sex)

male <- dplyr::filter(kbbh, kbbh$sex == 1) # neumerator
nm_male <- nrow(male)
percent_male <- round((nm_male/ dm) * 100,2)

female <- dplyr::filter(kbbh, kbbh$sex == 0) # neumerator
nm_female <- nrow(female)
percent_female <- round((nm_female/ dm) * 100,2)

# Type of Injury

#dm_tyi <- nrow(subset(kbbh, !is.na(tyi))) #denominator if excuding NA

kbbh$tyi <- as.numeric(kbbh$tyi)

blunt <- dplyr::filter(kbbh, kbbh$tyi == 0) # neumerator
nm_blunt <- nrow(blunt)
percent_blunt <- round((nm_blunt/ dm) * 100,2)

# Mode of Transport

kbbh$mot <- as.numeric(kbbh$mot)

kbbh$mot <- replace(kbbh$mot, kbbh$mot == 999, NA)

mot_pvt <- dplyr::filter(kbbh, kbbh$mot == 2) # neumerator
nm_pvt <- nrow(mot_pvt)
percent_mot_pvt <- round((nm_pvt/ dm) * 100,2)

mot_onfoot <- dplyr::filter(kbbh, kbbh$mot == 3) # neumerator
nm_onfoot <- nrow(mot_onfoot)
percent_mot_onfoot <- round((nm_onfoot/ dm) * 100,2)

```


```{r, results='hide'}

# Mechanism of Injury
moi.collapsed <- moi <- as.character(kbbh$moi)

## Method 2
## You can probably see how you could create a function to do some work for you
get_icd_range <- function(first.char, start.number, end.number) {
    icd.range <- paste0(
        paste0(
            "^",
            first.char,
            stringr::str_pad(start.number:end.number, 2, pad = "0"),
            "[0-9]?$"),
        collapse = "|")
    return (icd.range)
}

icd.ranges <- list(c("Transport accident" = get_icd_range("V", 0, 99)),
                   c("Fall" = get_icd_range("W", 0, 19)),
                   c("Animal bite" = get_icd_range("W", 50, 64)),
                   c("Assault" = paste0(get_icd_range("X", 85, 99), "|", get_icd_range("Y", 0, 9))))

for(icd.range in icd.ranges) moi.collapsed[grep(icd.range, moi)] <- names(icd.range)
moi.collapsed[!(moi.collapsed %in% sapply(icd.ranges, attr, "names"))] <- "Other"

## And if you want to put this back into the data as a factor then:
kbbh$moi.collapsed <- as.factor(moi.collapsed)

moi_assualt <- dplyr::filter(kbbh, kbbh$moi.collapsed == "Assault") # neumerator
nm_moi_assualt <- nrow(moi_assualt)
percent_moi_assualt <- round((nm_moi_assualt/ dm) * 100,2)

moi_tra <- dplyr::filter(kbbh, kbbh$moi.collapsed == "Transport accident") # neumerator
nm_moi_tra <- nrow(moi_tra)
percent_moi_tra <- round((nm_moi_tra/ dm) * 100,2)

moi_fall <- dplyr::filter(kbbh, kbbh$moi.collapsed == "Fall") # neumerator
nm_moi_fall <- nrow(moi_fall)
percent_moi_fall <- round((nm_moi_fall/ dm) * 100,2)

moi_anb <- dplyr::filter(kbbh, kbbh$moi.collapsed == "Animal bite") # neumerator
nm_moi_anb <- nrow(moi_anb)
percent_moi_anb <- round((nm_moi_anb/ dm) * 100,2)

moi_other <- dplyr::filter(kbbh, kbbh$moi.collapsed == "Other") # neumerator
nm_moi_other <- nrow(moi_other)
percent_moi_other <- round((nm_moi_other/ dm) * 100,2)

## Note that your performing similar operations many time
## above. That's usually a good indication that you can make your code
## more efficient and less prone to errors if you turn it into a loop
## of some sort

calculate_n_p_moi <- function(level.name) {
    moi.level <- dplyr::filter(kbbh, kbbh$moi.collapsed == level.name) # neumerator
    nm <- nrow(moi.level)
    percent <- round((nm/dm) * 100,2)
    return.list <- list(nm = nm, percent = percent)
    return (return.list)
}
moi.data <- lapply(setNames(nm = levels(kbbh$moi.collapsed)), calculate_n_p_moi)


# Trauma outcomes
kbbh$s24h <- as.numeric(kbbh$s24h)
kbbh$s30d <- as.numeric(kbbh$s30d)
kbbh$s6m <- as.numeric(kbbh$s6m)

kbbh$s24h <- replace(kbbh$s24h, kbbh$s24h == 999, NA)
kbbh$s30d <- replace(kbbh$s30d, kbbh$s30d == 999, NA)
kbbh$s6m <- replace(kbbh$s6m, kbbh$s6m == 999, NA)


## Alive at 6month
outcome_a <- dplyr::filter(kbbh, kbbh$s6m == "0") # neumerator
nm_outcome_a <- nrow(outcome_a)
percent_outcome_a <- round((nm_outcome_a/ dm) * 100,2)

## Status unknown/ Loss of followup
outcome_uk <- dplyr::filter(kbbh, is.na(kbbh$s6m)) # neumerator
nm_outcome_uk <- nrow(outcome_uk)
percent_outcome_uk <- round((nm_outcome_uk/ dm) * 100,2)

## Dead at 6month
outcome_d <- dplyr::filter(kbbh, kbbh$s6m == "1") # neumerator
nm_outcome_d <- nrow(outcome_d)
percent_outcome_d <- round((nm_outcome_d/ dm) * 100,2)

## See if you can apply the same function for calculating n and p that
## I defined above to make the outcome calculations more efficient

## Successful Followups
percent_followup <- percent_outcome_a+percent_outcome_d

```

In the TTRIS dataset, `r dm` patients were triaged green. The mean age of these patients was `r mean_age` (+-`r stddev_age`). `r percent_agebelow65`% of the patients were below 65 years of age. `r percent_male`% patients were male, while `r percent_female`% patients were female. `r percent_blunt`% patients had blunt injuries. Most common mode of arrival to the ED was through private vehicles<!-- Think about how you can get the name of this level without having to spell it out here --> (`r percent_mot_pvt`%). `r percent_mot_onfoot`% patients came to the ED on foot. Assualt<!-- You don't want to spell that out, you want to get that progamatically--> were the most common cause of injury `r percent_moi_assualt`% followed by transport accidents `r percent_moi_tra`% and fall `r percent_moi_fall`. Injury due to animal bites were `r percent_moi_anb`%.`r percent_followup`% patients were successfully followed up at 6months, of which `r nm_outcome_d` patients were reported dead. 

```{r}

kbbh$iss <- as.numeric(kbbh$iss)

# Categorizing ISS Score

kbbh$isspv[kbbh$iss >= 9 & kbbh$iss <= 15] = 'Moderate'
kbbh$isspv[kbbh$iss == 1 & kbbh$iss <= 8] = 'Minor'
kbbh$isspv[is.na(kbbh$iss)] = 'No injuries defined'

Table2 <- colPerc(xtabs(~kbbh$isspv+kbbh$moi.collapsed, data = kbbh))
kable(Table2, 
      caption = "Table 1. Percentage distribution of Injury severity among different Mechanism of Injury",
)

```

```{r}
# For calculating RTS

## Append a new variable 'GCS' to the dataset

kbbh$egcs <- as.numeric(kbbh$egcs)
kbbh$vgcs <- as.numeric(kbbh$vgcs)
kbbh$mgcs <- as.numeric(kbbh$mgcs)

gcs <- with(kbbh, egcs + vgcs + mgcs)

gcs <- replace(gcs, gcs > 15, NA)

gcspv <- gcs
gcspv[gcs >= 13] <- 4
gcspv[gcs >= 9 & gcs < 13] <- 3
gcspv[gcs >= 6 & gcs < 9] <- 2
gcspv[gcs >= 4 & gcs < 6] <- 1
gcspv[gcs < 4] <- 0

## RR parameter value for calculating RTS

kbbh$rr <- as.numeric(kbbh$rr)

rr <- with(kbbh,replace(rr, rr == 999, NA))

rrpv <- rr

rrpv[rr >= 10 & rr <= 29] = 4
rrpv[rr > 29] = 3
rrpv[rr >= 6 & rr <= 9] = 2
rrpv[rr >= 1 & rr <= 5] = 1
rrpv[rr == 0] = 0

## SBP parameter value for calculating RTS

kbbh$sbp <- as.numeric(kbbh$sbp)

sbp <- with(kbbh,replace(sbp, sbp == 999, NA))

sbppv <- sbp
sbppv[sbp > 89] <- 4
sbppv[sbp >=76 & sbp <= 89] <- 3
sbppv[sbp >= 50 & sbp <= 75] <- 2
sbppv[sbp >= 1 & sbp <= 49] <- 1
sbppv[sbp == 0] <- 0

## Formula for calculating RTS

## Why do you want to round it using 0 digits?
## So that I get a whole number.The decimals make a longer table.

kbbh$RTS = round(0.9368*as.numeric(gcspv) + 
                 0.7326*as.numeric(sbppv) + 
                 0.2908*as.numeric(rrpv), digits = 0)

Table2 <- rowPerc(xtabs(~kbbh$RTS, data = kbbh))
kable(Table2, 
      caption = "Table 1. RTS of patients",
)

```

## Conclusion

Among the patients triaged green, injury due to assualt and transport accidents were predominant.98.94% patients had an RTS score of 8. Patients with minor injuries and no injuries defined are upto 99.75%. 

## Analysis

```{r, warning=FALSE, results='hide'}

## Table of age group
tab1(kbbh$agegrp, main = "Distribution of patients by Age")

tab1(kbbh$sex, main = "Distribution of patients by Gender")

kbbh$tyi <- replace(kbbh$tyi, kbbh$tyi == 999, NA)

tab1(kbbh$tyi, main = "Distribution of type of injury")

kbbh$mot <- replace(kbbh$mot, kbbh$mot == 999, NA)

tab1(kbbh$mot, main = "Distribution of Mode of transport")

tab1(kbbh$RTS, main = "Distribution of RTS")

tab1(kbbh$hd, main = "Distribution of hospital disposition")

```



