---
title: "Profile of patients with RTI"
author: "Chandrika Verma"
date: "17/05/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
library(knitr)
library(psych)
library(data.table)
library(tidyverse)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

# Title
An observational study of adult trauma patients with history of road traffic injuries in urban India

# Introduction 
Increasing road traffic injuries is a major challenge for public health in India. The numbers of vehicles for transportation have been increasing every year. Age has been playing major role in road accidents<!-- This sentence is not clear. What do you mean by "Age has been playing major role"? Do you mean that young people are more often involved in road traffic crashes?-->. Though a road accident does not harm only physical body of human, it has impact on social, mental and economic impact. 

# Aim
To study the demographics of patients with road traffic injury presenting at a tertiary care hospital in Delhi, India<!-- You typically want to do more than just study something, as "study" is a very generic verb in this context. Can you think of other verbs that are more specific? Perhaps assess, determine etc.-->.

# Method

## Design
Proposed research is based on quantitative approach<!-- Use the active voice and be more specific. What design was it exactly? A cohort? Maybe you can write: "We conducted a prospective cohort study"-->". Proposed study also would be descriptive in nature<!-- I don't think you need this sentence-->.

## Study Setting
A tertiary care hospital emergency department. Participants aged 18 or above have been enrolled for the study those had history of road traffic accidents presented to emergency department.<!-- Use active voice and be direct: "We included patients aged at least 18 years who presented to the emergency department with road traffic injury"-->


## analysis of data<!-- Do you mean statistical analysis?-->
Data of 248 patients have been analyse with the help of R<!-- The final study sample is part of results, not methods-->. Some of the basic information has been withdrawn fron the current dataset.Some graphical repesentation has been done.<!-- Be more specific when you describe what you have done. Did you tabulate the demographics and present their summary statistics, for example?-->

```{r}

taftset <- read.csv("C:/Users/AMBIKA/Desktop/TAFT Analysis/taftset.csv") # You need to change the file name here.

summary(taftset$sex) 

prop.table(table(taftset$sex)) # It shows 79 percent are male and 21 percent female patients out of 248 patients.

hist(taftset$"sex", main = "Sex Distribution of Patients", xlab = "Sex", border = "blue", col="green", las=2, breaks =6)


hist(taftset$"age", main = "Age Distribution of Patients", xlab = "Age group", border = "blue",col="green", xlim=c(18,100), las=1, breaks =5)


counts<- table(taftset$tyi)

barplot(counts, main = "Type of Injury Distribution",
        xlab = "Type of Injury") # Majority of injury reported as blunt and a very less proportion reported penetrating.


```

# Result 
Based on data analysis result would be explained here though analysis has not been completed once this section would be complete discription will be given.

# Conclusion 
Conclusion would be drawn with the help of result.
