---
title: "Profile of patients with RTI"
author: "Chandrika Verma"
date: "17/05/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
```
# Abstract

# Title
An observational study of adult trauma patients with the history of road traffic injuries in urban India

# Introduction 
Increasing road traffic injuries have been a major challenge issue to public health in India. Numbers of vehicles for transportation have been increasing every year. Age has been playing major role in road accidents. Though a road accident does not harm only physical body of human, it has impact on social, mental and economic impact. 

# Aim
To study the demographic analysis of road accidents presented at tertiary care hospital in Delhi.

# Design
Proposed research is based on quantitative approach and retrospective in nature. Tertiary care hospital emergency department Delhi is the site of this study. Participants aged 18 or above have been enrolled for the study those had history of road traffic accidents presented to emergency department.

# Method
Data of 248 patients have been analysed with the help of R. Some of the basic information has been drawn.Some graphical representation has been done.

```{r}
## Load packages
library(knitr)
library(psych)
library(data.table)
library(tidyverse)
library(ggplot2)
library(rio)
## Let whoever runs this script, including you, put the path to the
## direcctory where the data is in an option called data.dir. You do
## that by running options(data.dir = "<path to data directory>"). You
## need to replace the <path to data directory> with the actual path.
## Data Directory 
data.dir <- getOption("C:/Users/CP Verma/Desktop/Git R abstract/abstract-challenge")
taft <- import(file.path("C:/Users/CP Verma/Desktop/Git R abstract", "taft-dataset-9825-20200608232208.csv")) %>% as_tibble()

## Age distribution 
summary(taft$age)
agebreaks <- c(17,25,45,65,80)
agelabels <- c("18-24","25-44","45-64","65+")
agegrup <-cut(taft$age,
                breaks = agebreaks,
                right = FALSE,
               labels = agelabels)
table1 <- table(agegrup)
table1


## Sex Distribution

summary(taft$sex) 

## Get percentage males

p.males <- round(mean(taft$sex == 1, na.rm = TRUE) * 100)
p.males

## Get percentage females
p.females <- round(mean(taft$sex == 0, na.rm = TRUE) * 100)
p.females

## ggplot2 is a good package for visualizing data. If you want to put
## a plot into your results, do this:

taft$sex.cat <- factor(taft$sex, levels = c(0, 1), labels = c("Female", "Male"))
plot.sex <- ggplot(taft, aes(sex.cat, fill = sex.cat)) +
    geom_bar() +
    ylab("Count") +
    xlab("Sex") +
    theme(legend.position = "none") 
plot.sex

## Type of Injury


## Mechanism of Injury


## Mode of Transportation
ambulance <- taft$mot == 0
n.ambulance <- sum(ambulance)
n.ambulance
p.ambulance <- round(mean(ambulance)*100)
p.ambulance


police <- taft$mot == 1
n.police <- sum(police)
n.police
p.police <- round(mean(police)*100)
p.police

on_feet <- taft$mot == 3
n.on_feet <- sum(on_feet)
n.on_feet
p.on_feet <- round(mean(on_feet)*100)
p.on_feet

unknown <- taft$mot == 999
n.unknown <- sum(unknown)
n.unknown
p.unknown <- round(mean(unknown)*100)
p.unknown


## Transfer Status of Patients

trans_status <- taft$tran == 1
n.trans_status <- sum(trans_status)
p.trans_status <- round(mean(trans_status)*100)
p.trans_status


```


# Result
The demography of the road accident victims `r p.males` percent are male and remaining `r 100 - p.males` percent are female (Figure 1). According to their age distribution it shows majority of the road accident victims are among age-group 18-40 years and majority of them  presented emergency department with blunt trauma. (Though I have not checked missing value if found some missing data from data set will try to explain my analysis on that basis after excluding missing items.  )

```{r figure-1, include = TRUE, fig.height = 3, fig.width = 3, fig.align = "center"}
plot.sex
```

# Conclusion 
Conclusion would be drawn with the help of result.
