---
title: "Profile of patients with RTI"
author: "Chandrika Verma"
date: "17/05/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
library(knitr)
library(psych)
library(data.table)
library(tidyverse)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

# Title
An observational study of adult trauma patients with the history of road traffic injuries in urban India

# Introduction 
Increasing road traffic injuries have been a major challenge issue to public health in India. Numbers of vehicles for transportation have been increasing every year. Age has been playing major role in road accidents. Though a road accident does not harm only physical body of human, it has impact on social, mental and economic impact. 

# Aim
To study the demographic analysis of road accidents presented at tertiary care hospital in Delhi.

# Method

## Research Desigh
 Proposed research is based on quantitative approach and retrospective in nature.
Tertiary care hospital emergency department Delhi is the site of this study. Participants aged 18 or above have been enrolled for the study those had history of road traffic accidents presented to emergency department.


## analysis of data
Data of 248 patients have been analyse with the help of R. Some of the basic information has been drawn.Some graphical representation has been done.

```{r}

taft.dataset.9825.20200608232208 <- read.csv("C:/Users/acer/Desktop/TAFT Analysis/taft-dataset-9825-20200608232208.csv")



summary(taft.dataset.9825.20200608232208$sex) 

prop.table(table(taft.dataset.9825.20200608232208$sex)) # It shows 79 percent are male and 21 percent female patients out of 248 patients.

hist(taft.dataset.9825.20200608232208$"sex", main = "Sex Distribution of Patients", xlab = "Sex", border = "blue", col="green", las=2, breaks =6)


hist(taft.dataset.9825.20200608232208$"age", main = "Age Distribution of Patients", xlab = "Age group", border = "blue",col="green", xlim=c(18,100), las=1, breaks =5)


counts<- table(taft.dataset.9825.20200608232208$tyi)

barplot(counts, main = "Type of Injury Distribution",
        xlab = "Type of Injury") # Majority of injury reported as blunt and a very less proportion reported penetrating.


```

# Result
The demography of the road accident victims 79 percent are male and rest 21 percent are female.According to their age distribution it shows majority of the road accident victims are among age-group 18-40 years and majority of them  presented emergency department with blunt trauma. 

(Though I have not checked missing value if found some missing data from data set will try to explain my analysis on that basis after excluding missing items.  )

# Conclusion 
Conclusion would be drawn with the help of result and explain here.
