---
title: Profile of Patients with Thoracic Trauma at Tertiary Care Hospital in Delhi
author: "Deepak Kumar"
date: "17/05/2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

# Introduction 
  A chest injury, also known as chest trauma, is any form of physical injury to the chest including the heart, ribs and lungs. Chest injuries account for 25% of all deaths from traumatic injury. Typically chest injuries are caused by blunt mechanisms such as road traffic accidents or penetrating mechanisms such as stabbing. Blunt trauma refers to mechanisms causing increased in thoracic pressure such as injuries most common cause of thoracic trauma), falls and other causes of road traffic injuries.In the current study only thoracic trauma patients will be include. 

# Aim
     To study the a demographic factors associated inthe thoracic trauma patients and their outcomes.

# Design
In our study 18 & above age group persons are enrolled.

# Method
 Present study would be quantitative in nature.
```{r}
TAFTDATA <- read.csv("D:/taft r/TAFTDATA.csv")
summary(TAFTDATA$age)
```
```{r}
hist(TAFTDATA$tyi,col = 'blue')
```
```{r}
ggplot(TAFTDATA, aes(x= age)) + geom_bar()

```
```{r}
prop.table(table (TAFTDATA$)
```
```{r}
print(seq(1,20))



```

```{r}
mystring<- "this is my output for analysis"
print(mystring)
```
 


  * ** This histogram shows type of injury.(0=Blunt, 1= Penetrating)**
  
# Result
  The TAFT study data analysis of 249 patients at Lok Nayak Hospital, New Delhi shows around 99% patients having blunt injury, most of these patients are injured in road accidents. Its shows the main reason of thoracic injury is blunt trauma.
    
# Conclusion
The Present study has concluded majority of thoracic trauma is due blunt injury among road accident patients.