---
title: "ABDOMINAL TRAUMA AND SURVIVAL"
author: "KUNAL NATH"
date: "June 2, 2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INTRODUCTION

 Abdominal injuries are common, costly, deadly, and a source of considerable uncertainty common in adult  trauma. In some circumstances, laparoscopy offers a diagnostic and therapeutic alternative with less morbidity than standard exploration, but in some specific cases morbidity is high .
 
# AIM

TO check the survival rate of abdominal trauma cases.

# Desing

Most of the paitent with polytrauma less chance of having abdominal trauma.

1.so selection is very much requered.

2.finding proper history

3.specify injury from medical record

 
# Methods

Most of the abdominal trauma paitent having polytrauma.
paitent inclution critriea 18 to 100 (age) so adult paitent only involve in this study.

Selection by mechanism of injuri.with medical record

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
