---
title: "Adult age and Red triage"
author: "TAMAL"
date: "06/06/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
It is important to triage trauma patients because proper triaging indicate level of care required for the patients. The red triage level is used for very serious patients where green triage is for minor injury.So triaging is very important part of trauma care system.Present data shows that trauma increasing in adult age rather than senior adult.

## Aim
Aim of the study is To characterise trauma patients triaged red with special reference to their age distribution.

## Study design
Data patients triage red were extracted from TTRIS data set of centre 1515 then retrospective analysis of a prospective cohort of trauma patients with age division.


## Methods
Data of triaged red patients extract out from the data set of TTRIS study project. A retrospective cohort analysis of the data will performed using R with respect to age group. 

## Result
```{r}
TTRIS_1515 <- read.csv("C:/Users/Tamal/OneDrive/Documents/TTRIS_1515.csv")
summary(TTRIS_1515$tyi)
summary(TTRIS_1515$age)
hist(TTRIS_1515$age, col='grey',xlab='Age', main='Age of patients')
n.patients <- nrow(TTRIS_1515)
mean.age <- mean(TTRIS_1515$age, na.rm = TRUE)
summary(TTRIS_1515$tc)
hist(TTRIS_1515$tc, col='grey',xlab='tc', main='Triage of patients')
summary(TTRIS_1515$tc == 3)
table(TTRIS_1515$tc == 3)
red <- subset(x=TTRIS_1515, subset = TTRIS_1515$tc == 3)
table(red$tc)
fd_red <- subset(x=red, subset = red$age <61)
table(fd_red$doa)
table(fd_red$doa == 0)
```
table(fd_red$age)
table(fd_red$sex == 1)
mean(fd_red$sex == 1)*100
table(fd_red$sex == 0)
mean(fd_red$sex == 0)*100
table(fd_red$moi)
death <- subset(x=fd_red, subset = fd_red$hd == 1)
table(death$hd)
mean(TTRIS_1515$tc == 3)*100
mean(red$hd == 1)*100
mean(red$age <61)*100
```

Total number patients available for analysis n= 5618, triage red category patients are 272 under this category so 4% patients(red dataset) are triage red, adult age patients under this triage are 219 percentage is 80(fd_red data set),Admitted patients are 183 patients non admitted 36 patients, male patients are 187 i.e 85.38% and female patients are 32 i.e 14.61%. and Number of patients death occured in this triage category are 158, percentage is 74.26. 

##Conclusion

So from this data set it can conclude that death rate of of triage red is very much high in this perticular hospital.