---
output: pdf_document
---

```{r create-example-data, echo = FALSE, include = FALSE}
## These first lines are just to create some example data. You don't
## have to do this when you have your own data
data(mtcars)
mtcars[1, 4] <- NA
mtcars[3, 2] <- NA
mtcars[7, 4] <- NA
mtcars[20, 6] <- NA
mtcars[31, 8] <- NA
study.data <- mtcars
```


Results
=======
```{r load-packages, echo = FALSE, include = FALSE}
library(rmarkdown)
library(tableone)
```

```{r summarize-missing-data, echo = FALSE, include = FALSE}
## If you are going to do a complete case analysis it makes sense to
## implement it as early as possible. But first you probably want a
## table that shows how much missing you have in each variable
missing.data.list <- lapply(study.data, function(column) {
    n.na <- sum(is.na(column))
    p.na <- round((n.na/length(column)) * 100)
    missing <- data.frame("Count" = n.na, "Percentage" = p.na)
    return (missing)
})
missing.data <- do.call(rbind, missing.data.list)
## You can then use this table to identify which variable had the most
## missing data
maximum.missing <- rownames(missing.data)[missing.data$Count == max(missing.data$Count)]
## Now let's assign the number and percentages of incomplete cases to
## an object
complete.index <- complete.cases(study.data)
n.incomplete <- sum(!complete.index) # Note the use of the exclamation mark !
p.incomplete <- round((n.incomplete/nrow(study.data)) * 100)
```

There was `r nrow(study.data)` observations in the study data. Out of these, `r n.incomplete` (`r p.incomplete` %) observations were incomplete. Table 1 shows how the missing data was distributed. The variable with most missing values was `r maximum.missing` with `r missing.data[maximum.missing, "Count"]` (`r missing.data[maximum.missing, "Percentage"]` %) values missing.

```{r missing-data-table, echo = FALSE}
knitr::kable(missing.data, caption = "Count and percentage of missing values per variable")
```

After excluding the observations with missing values the study sample included `r nrow(study.data) - n.incomplete` observations. The supplementary material includes a comparisons of complete and incomplete cases.

```{r create-study-sample, echo = FALSE, include = FALSE}
## Remove incomplete observations
study.sample <- study.data[complete.index, ]
## Then you go on to summarize your study sample here
```

Supplementary material
======================

```{r complete-incomplete-comparison, echo = FALSE, include = FALSE}
## Let's create a table that contrasts complete and incomplete observations
study.data$complete <- factor(complete.index,
                              levels = c(FALSE, TRUE),
                              labels = c("Incomplete", "Complete"))
variables <- colnames(study.data)[-grep("complete", colnames(study.data))]
comparison.table <- tableone:::print.TableOne(CreateTableOne(vars = variables,
                                         data = study.data,
                                         strata = "complete",
                                         test = FALSE),
                                         printToggle = FALSE)
```

```{r comparison-table, echo = FALSE}
knitr::kable(comparison.table, caption = "Comparison of complete and incomplete observations")
```

